---
title: "Sri Lanka Dengue Dashboard"
format:
  dashboard:
    css: styles.css
    orientation: rows
    mobile: true
editor: source
---

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(denguedatahub)
library(ceylon)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(here)
library(stringdist)
library(sf)
library(viridis)
library(cartogram)
library(ISOweek)
library(lubridate)
library(xts)
library(dygraphs)
library(crosstalk)
library(leaflet)
library(htmltools)
library(plotly)
library(shiny)
library(RColorBrewer)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(scales) 
library(GGally)
library(DCCA)
library(patchwork)
```

```{r}
#| label: load-data
#| message: false
#| warning: false
data(srilanka_weekly_data)

weekly_2026 <- read_csv(here("data","sri_lanka_weekly_dengue_cases_2026.csv"))
weekly_2026$year        <- as.numeric(weekly_2026$year)
weekly_2026$week        <- as.numeric(weekly_2026$week)
weekly_2026$start.date  <- as.Date(weekly_2026$start.date)
weekly_2026$district    <- as.character(weekly_2026$district)
weekly_2026$cases       <- as.numeric(weekly_2026$cases)
weekly_2026 <- weekly_2026 |>
  mutate(
    district = if_else(district == "Kalmune", "Ampara", district)
  ) |>
  group_by(year, week, start.date, end.date, district) |>
  summarise(cases = sum(cases), .groups = "drop")


Global <- read_excel(here("data", "dengue-global-data.xlsx"))
Global$date      <- as.Date(Global$date )
Global$cases     <- as.numeric(Global$cases)
Global$country   <- as.character(Global$country )
Global$deaths    <- as.numeric(Global$deaths)

weekly_climate <- read_csv(here("data", "weekly_climate.csv"))

last_week_comparisson <- read_excel("data/last_week_comparisson.xlsx")
last_week_comparisson$district <- as.character(last_week_comparisson$district)
last_week_comparisson$`2025_26th`<- as.numeric(last_week_comparisson$`2025_26th`)
last_week_comparisson$`2026_26th`<- as.numeric(last_week_comparisson$`2026_26th`)
last_week_comparisson$deaths2026 <- as.numeric(last_week_comparisson$deaths2026)
last_week_comparisson <- last_week_comparisson |>
  mutate(
    district = if_else(district == "Kalmune", "Ampara", district)
  ) |>
  group_by(district) |>
  summarise(
    `2025_26th` = sum(`2025_26th`, na.rm = TRUE),
    `2026_26th` = sum(`2026_26th`, na.rm = TRUE),
    deaths2026 = sum(deaths2026, na.rm = TRUE),
    .groups = "drop"
  )
```

# Overview

::: yellow-box
**Latest Dengue Surveillance Summary**\
**Last Updated:** Tuesday, July 07, 2026\| **Latest local data available for:** June 28, 2026
:::

## Row {height="20%"}

```{r}
#| content: valuebox
#| title: "Total Cases in 2026"
#| color: "#5DADE2"

format(
  sum(weekly_2026$cases, na.rm = TRUE),
  big.mark = ",",
  scientific = FALSE
)
```

```{r}
#| content: valuebox
#| title: "Total Deaths in 2026"
#| color: "#DC2626"

format(
  sum(last_week_comparisson$deaths2026, na.rm = TRUE),
  big.mark = ",",
  scientific = FALSE
)
```

```{r}
#| content: valuebox
#| title: "Average Weekly Cases in 2026"
#| color: "#7570b3"

format(
  round(sum(weekly_2026$cases, na.rm = TRUE)/26),
  big.mark = ",",
  scientific = FALSE
)

```

```{r}
#| content: valuebox
#| title: "Cases in Last Week"
#| color: "#FFFF00"

format(
  weekly_2026 |>
  filter(week == 26) |>
  summarise(total_cases = sum(cases, na.rm = TRUE)) |>
  pull(total_cases),
  big.mark = ",",
  scientific = FALSE
)
```

```{r}
#| content: valuebox
#| title: "Highest Risk District in 2026"
#| color: "#1b9e77"

  weekly_2026 |>
  group_by(district) |>
  summarise(total_cases = sum(cases, na.rm = TRUE)) |>
  arrange(desc(total_cases)) |>
  slice(1) |>
  pull(district)

```

## Row {height="65%"}

### Column {width="40%"}

```{r}
#| label: map1
#| echo: false
#| message: false
#| warning: false

district_df <- weekly_2026 |>
  group_by(district) |>       
  summarise(Count = sum(cases), .groups = "drop")  

data(district)
district_df <- district_df |>
  mutate(district_clean = toupper(trimws(district)))

district_sf <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df <- district_df |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf$DISTRICT_clean[
      amatch(district_clean, district_sf$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2 <- as.data.frame(district_df)

map_data <- left_join(
  district_sf,
  district_df2,
  by = c("DISTRICT_clean" = "DISTRICT_match")
) |>
drop_na(Count)  


map_data <- map_data |>
  mutate(
    Cases_Category = case_when(
      Count >= 0 & Count <= 30       ~ "0-30",
      Count >= 31 & Count <= 60      ~ "31-60",
      Count >= 61 & Count <= 100     ~ "61-100",
      Count >= 101 & Count <= 150    ~ "101-150",
      Count >= 151 & Count <= 250    ~ "151-250",
      Count >= 251 & Count <= 500    ~ "251-500",
      Count >= 501 & Count <= 1000   ~ "501-1000",
      Count >= 1001 & Count <= 3000  ~ "1001-3000",
      Count > 3000                   ~ "Above 3000",
      TRUE                           ~ NA_character_
    )
  )

map_data$Cases_Category <- factor(
  map_data$Cases_Category,
  levels = c(
    "0-30",
    "31-60",
    "61-100",
    "101-150",
    "151-250",
    "251-500",
    "501-1000",
    "1001-3000",
    "Above 3000"
  )
)


g1 <- ggplot(map_data) +
  geom_sf(
  aes(
    fill = Cases_Category,
    text = paste0(
      "District: ", DISTRICT,
      "<br>Total Cases: ", Count
    )
  ),
    color = "white",
    linewidth = 0.2
  ) +
  coord_sf(expand = FALSE) +
  scale_fill_manual(
    values = c(
      "0-30"       = viridis(9)[9],
      "31-60"      = viridis(9)[8],
      "61-100"     = viridis(9)[7],
      "101-150"    = viridis(9)[6],
      "151-250"    = viridis(9)[5],
      "251-500"    = viridis(9)[4],
      "501-1000"   = viridis(9)[3],
      "1001-3000"  = viridis(9)[2],
      "Above 3000" = viridis(9)[1]
    ),
    name = "Cases",
    drop = FALSE
) +
  labs(
    title = "Total Dengue Cases in Sri Lanka (2026)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    ),
    plot.margin = margin(5, 5, 5, 5)
  )


ggplotly(g1, tooltip = "text") |> 
  layout(
    autosize = TRUE
  )

```

### Column {width="40%"}

```{r}
#| label: map2
#| echo: false
#| message: false
#| warning: false


district_df_death <- last_week_comparisson |>
  select(district, deaths2026)

data(district)
district_df_death <- district_df_death |>
  mutate(district_clean = toupper(trimws(district)))

district_sf_death <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df_death <- district_df_death |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf_death$DISTRICT_clean[
      amatch(district_clean, district_sf_death$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2_death <- as.data.frame(district_df_death)

map_data_death <- left_join(
  district_sf_death,
  district_df2_death,
  by = c("DISTRICT_clean" = "DISTRICT_match")
) |>
drop_na(deaths2026)  


map_data_death <- map_data_death |>
  mutate(
    death_cat = case_when(
      deaths2026 == 0               ~ "0",
      deaths2026 >= 1 & deaths2026 <= 2 ~ "1–2",
      deaths2026 >= 3 & deaths2026 <= 5 ~ "3–5",
      deaths2026 >= 6 & deaths2026 <= 9 ~ "6–9",
      deaths2026 >= 10              ~ "10+",
      TRUE                           ~ NA_character_
    )
  )


map_data_death$death_cat <- factor(
  map_data_death$death_cat,
  levels = c("0", "1–2", "3–5", "6–9", "10+")
)


g2 <- ggplot(map_data_death) +
  geom_sf(aes(fill = death_cat,
              text = paste0(
        "District: ", district,
        "<br>Cumulative Deaths: ", deaths2026
      )),
          colour = "black",
          linewidth = 0.4) +
  coord_sf(expand = FALSE) +
  scale_fill_manual(
  values = c(
    "0"   = "#D9D9D9",  
    "1–2" = "#F8B4B4",  
    "3–5" = "#F46D6D",  
    "6–9" = "#D73027",  
    "10+" = "#7F0000"   
  ),
  name = "Deaths",
  drop = FALSE
) +
  labs(
    title = "Cumulative Deaths in Sri Lanka (2026)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    ),
    plot.margin = margin(5, 5, 5, 5)
  )

ggplotly(g2, tooltip = "text") |> 
  layout(
    autosize = TRUE
  )
```

### Column {width="40%"}

```{r}
#| label: TS1
#| echo: false
#| message: false
#| warning: false


srilanka_weekly_dengue_past <- srilanka_weekly_data |>
  mutate(
    start.date = mdy(start.date),
    end.date = mdy(end.date)
  )

srilanka_weekly_dengue_past <- srilanka_weekly_dengue_past |>
  mutate(
    Week_start = floor_date(start.date, unit = "week", week_start = 1)  # Monday as start of week
  ) 

weekly_past_cases_by_date <- srilanka_weekly_dengue_past |>
  group_by(start.date) |>
  summarise(weekly_cases = sum(cases, na.rm = TRUE)) |>
  arrange(start.date)



weekly_cases_by_date <- weekly_2026 |>
  group_by(start.date) |>
  summarise(weekly_cases = sum(cases, na.rm = TRUE)) |>
  arrange(start.date)



past_data <- weekly_past_cases_by_date %>%
  mutate(
    year = year(start.date),
    week = isoweek(start.date)   
  ) %>%
  filter(year %in% c(2024, 2025))


data_2026 <- weekly_cases_by_date %>%
  mutate(
    year = 2026,
    week = isoweek(start.date)
  )


weekly_all <- bind_rows(past_data, data_2026) %>%
  mutate(year = factor(year))

weekly_all$week[103:128] <- 1:26



g3 <- ggplot(
  weekly_all,
  aes(
    x = week,
    y = weekly_cases,
    color = year,
    group = year
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 0.7) +
  scale_color_viridis_d(option = "D", end = 0.9, name = "Year") +
  scale_x_continuous(
    breaks = seq(1, 52, by = 2),
    limits = c(1, 52)
  ) +
  labs(
    title = "Weekly Dengue Cases (2024–2026)",
    x = "Week Number",
    y = "Cases"
  ) +
  coord_cartesian(ylim = c(0, 6000)) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 10, hjust = 0.5),
    axis.title = element_text(size = 13),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 7),
    axis.text.y = element_text(size = 11),
    legend.title = element_text(face = "bold")
  )

ggplotly(g3) |>
  layout(autosize = TRUE)
```

# Last Week Dengue Situation Analysis

## Row {height="100%"}

### Column {width="50%"}

```{r}
#| label: Lweeek1
#| echo: false
#| message: false
#| warning: false



district_df_2026 <- last_week_comparisson |>
  select(district, `2026_26th`)

data(district)
district_df_2026 <- district_df_2026 |>
  mutate(district_clean = toupper(trimws(district)))

district_sf_2026 <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df_2026 <- district_df_2026 |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf_2026$DISTRICT_clean[
      amatch(district_clean, district_sf_2026$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2_2026 <- as.data.frame(district_df_2026)

map_data_2026 <- left_join(
  district_sf_2026,
  district_df2_2026,
  by = c("DISTRICT_clean" = "DISTRICT_match")
) |>
drop_na(`2026_26th`)  


map_data_2026 <- map_data_2026 |>
  mutate(
    case2026_cat = case_when(
      
      `2026_26th` == 0                         ~ "0",
      
      `2026_26th` >= 1 & `2026_26th` <= 10     ~ "1-10",
      
      `2026_26th` >= 11 & `2026_26th` <= 30    ~ "11-30",
      
      `2026_26th` >= 31 & `2026_26th` <= 60    ~ "31-60",
      
      `2026_26th` >= 61 & `2026_26th` <= 100   ~ "61-100",
      
      `2026_26th` >= 101 & `2026_26th` <= 150  ~ "101-150",
      
      `2026_26th` >= 151 & `2026_26th` <= 250  ~ "151-250",
      
      `2026_26th` >= 250                       ~ "250+",
      
      TRUE ~ NA_character_
    )
  )



map_data_2026$case2026_cat <- factor(
  map_data_2026$case2026_cat,
  levels = c(
    "0",
    "1-10",
    "11-30",
    "31-60",
    "61-100",
    "101-150",
    "151-250",
    "250+"
  )
)


p1 <- ggplot(map_data_2026) +
  geom_sf(aes(fill = case2026_cat),
          colour = "black",
          linewidth = 0.4) +
  geom_sf_text(
    aes(label = `2026_26th`),
    size = 3.25,
    color = "black",
    fontface = "bold"
  ) +
  coord_sf(expand = FALSE) +
 scale_fill_manual(
    values = c(
      "0"       = alpha("#BDBDBD", 0.5),
      "1-10"    = alpha("#FFF3E0", 0.5),
      "11-30"   = alpha("#FFE0B2", 0.5),
      "31-60"   = alpha("#FFCC80", 0.5),
      "61-100"  = alpha("#FFB74D", 0.5),
      "101-150" = alpha("#FB8C00", 0.5),
      "151-250" = alpha("#EF6C00", 0.5),
      "250+"    = alpha("#B33A00", 0.5)
    ),
    name = "Cases",
    drop = FALSE
  ) +
  labs(
    title = "Cases in Last Week (26th Week)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    ),
    plot.margin = margin(5, 5, 5, 5)
  )

p1
```

### Column {width="50%"}

```{r}
#| label: Lweeek2
#| echo: false
#| message: false
#| warning: false



district_df_diff <- last_week_comparisson |>
  mutate(
    diff = `2026_26th` - `2025_26th`
  ) |>
  select(district, diff)

data(district)
district_df_diff <- district_df_diff |>
  mutate(district_clean = toupper(trimws(district)))

district_sf_diff <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df_diff <- district_df_diff |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf_diff$DISTRICT_clean[
      amatch(district_clean, district_sf_diff$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2_diff <- as.data.frame(district_df_diff)

map_data_diff <- left_join(
  district_sf_diff,
  district_df2_diff,
  by = c("DISTRICT_clean" = "DISTRICT_match")
) |>
drop_na(diff)

map_data_diff <- map_data_diff |>
  mutate(
    casediff_cat = case_when(
      
      diff == 0                         ~ "0",
      
      diff >= 1 & diff <= 10     ~ "1-10",
      
      diff >= 11 & diff <= 30    ~ "11-30",
      
      diff >= 31 & diff <= 60    ~ "31-60",
      
     diff >= 61 & diff <= 100   ~ "61-100",
      
      diff >= 101 & diff <= 150  ~ "101-150",
      
      diff >= 151 & diff <= 250  ~ "151-250",
      
      diff >= 250                       ~ "250+",
      
      TRUE ~ NA_character_
    )
  )



map_data_diff$casediff_cat <- factor(
  map_data_diff$casediff_cat,
  levels = c(
    "0",
    "1-10",
    "11-30",
    "31-60",
    "61-100",
    "101-150",
    "151-250",
    "250+"
  )
)

map_data_diff <- map_data_diff |>
  mutate(
    diff_label = ifelse(
      diff > 0,
      paste0("+", diff),
      as.character(diff)
    )
  )

p2 <- ggplot(map_data_diff) +
  geom_sf(aes(fill = casediff_cat),
          colour = "black",
          linewidth = 0.4) +
  geom_sf_text(
    aes(label = diff_label),
    size = 3.25,
    color = "black",
    fontface = "bold"
  ) +
  coord_sf(expand = FALSE) +
  scale_fill_manual(
    values = c(
      "0"       = alpha("#BDBDBD", 0.5),
      "1-10"    = alpha("#FFF3E0", 0.5),
      "11-30"   = alpha("#FFE0B2", 0.5),
      "31-60"   = alpha("#FFCC80", 0.5),
      "61-100"  = alpha("#FFB74D", 0.5),
      "101-150" = alpha("#FB8C00", 0.5),
      "151-250" = alpha("#EF6C00", 0.5),
      "250+"    = alpha("#B33A00", 0.5)
    ),
    name = "Cases",
    drop = FALSE
  ) +
  labs(
    title = "Additional Cases in 26th Week Compared to that Week in 2025"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.key = element_rect(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    ),
    plot.margin = margin(5, 5, 5, 5)
  )  
    
  p2

```

# District-wise Time Series Trends

## Column {.tabset}

### Overview of District Trends

```{r}
#| label: TS2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20

 
srilanka_weekly_df <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  )

g4 <- ggplot(srilanka_weekly_df, aes(x = week_start, y = cases)) +
  geom_line(color = "darkblue", linewidth = 0.8) +
  facet_wrap(vars(district), scales = "free_y", ncol = 4) + 
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    x = "Year",
    y = "Weekly dengue cases"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 32, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28)
  )

g4
```

### Explore Individual District Trends

```{r}
#| label: TS3
#| echo: false
#| warning: false
#| message: false


srilanka_weekly <- srilanka_weekly_data |>
  mutate(start.date = mdy(start.date)) |>
  arrange(start.date)

# --- Colombo ---
colombo_data <- srilanka_weekly |> filter(district == "Colombo")
colombo_xts  <- xts(colombo_data$cases, order.by = colombo_data$start.date)
colnames(colombo_xts) <- "Cases"

g_colombo <- dygraph(colombo_xts, main = "Colombo: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Gampaha ---
gampaha_data <- srilanka_weekly |> filter(district == "Gampaha")
gampaha_xts  <- xts(gampaha_data$cases, order.by = gampaha_data$start.date)
colnames(gampaha_xts) <- "Cases"

g_gampaha <- dygraph(gampaha_xts, main = "Gampaha: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalutara ---
kalutara_data <- srilanka_weekly |> filter(district == "Kalutara")
kalutara_xts  <- xts(kalutara_data$cases, order.by = kalutara_data$start.date)
colnames(kalutara_xts) <- "Cases"

g_kalutara <- dygraph(kalutara_xts, main = "Kalutara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kandy ---
kandy_data <- srilanka_weekly |> filter(district == "Kandy")
kandy_xts  <- xts(kandy_data$cases, order.by = kandy_data$start.date)
colnames(kandy_xts) <- "Cases"

g_kandy <- dygraph(kandy_xts, main = "Kandy: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matale ---
matale_data <- srilanka_weekly |> filter(district == "Matale")
matale_xts  <- xts(matale_data$cases, order.by = matale_data$start.date)
colnames(matale_xts) <- "Cases"

g_matale <- dygraph(matale_xts, main = "Matale: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Nuwara Eliya ---
nuwara_eliya_data <- srilanka_weekly |> filter(district == "NuwaraEliya")
nuwara_eliya_xts  <- xts(nuwara_eliya_data$cases, order.by = nuwara_eliya_data$start.date)
colnames(nuwara_eliya_xts) <- "Cases"

g_nuwara_eliya <- dygraph(nuwara_eliya_xts, main = "Nuwara Eliya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Galle ---
galle_data <- srilanka_weekly |> filter(district == "Galle")
galle_xts  <- xts(galle_data$cases, order.by = galle_data$start.date)
colnames(galle_xts) <- "Cases"

g_galle <- dygraph(galle_xts, main = "Galle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matara ---
matara_data <- srilanka_weekly |> filter(district == "Matara")
matara_xts  <- xts(matara_data$cases, order.by = matara_data$start.date)
colnames(matara_xts) <- "Cases"

g_matara <- dygraph(matara_xts, main = "Matara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Hambantota ---
hambantota_data <- srilanka_weekly |> filter(district == "Hambanthota")
hambantota_xts  <- xts(hambantota_data$cases, order.by = hambantota_data$start.date)
colnames(hambantota_xts) <- "Cases"

g_hambantota <- dygraph(hambantota_xts, main = "Hambanthota: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Jaffna ---
jaffna_data <- srilanka_weekly |> filter(district == "Jaffna")
jaffna_xts  <- xts(jaffna_data$cases, order.by = jaffna_data$start.date)
colnames(jaffna_xts) <- "Cases"

g_jaffna <- dygraph(jaffna_xts, main = "Jaffna: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kilinochchi ---
kilinochchi_data <- srilanka_weekly |> filter(district == "Kilinochchi")
kilinochchi_xts  <- xts(kilinochchi_data$cases, order.by = kilinochchi_data$start.date)
colnames(kilinochchi_xts) <- "Cases"

g_kilinochchi <- dygraph(kilinochchi_xts, main = "Kilinochchi: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mannar ---
mannar_data <- srilanka_weekly |> filter(district == "Mannar")
mannar_xts  <- xts(mannar_data$cases, order.by = mannar_data$start.date)
colnames(mannar_xts) <- "Cases"

g_mannar <- dygraph(mannar_xts, main = "Mannar: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Vavuniya ---
vavuniya_data <- srilanka_weekly |> filter(district == "Vavuniya")
vavuniya_xts  <- xts(vavuniya_data$cases, order.by = vavuniya_data$start.date)
colnames(vavuniya_xts) <- "Cases"

g_vavuniya <- dygraph(vavuniya_xts, main = "Vavuniya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mullaitivu ---
mullaitivu_data <- srilanka_weekly |> filter(district == "Mullaitivu")
mullaitivu_xts  <- xts(mullaitivu_data$cases, order.by = mullaitivu_data$start.date)
colnames(mullaitivu_xts) <- "Cases"

g_mullaitivu <- dygraph(mullaitivu_xts, main = "Mullaitivu: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Batticaloa ---
batticaloa_data <- srilanka_weekly |> filter(district == "Batticaloa")
batticaloa_xts  <- xts(batticaloa_data$cases, order.by = batticaloa_data$start.date)
colnames(batticaloa_xts) <- "Cases"

g_batticaloa <- dygraph(batticaloa_xts, main = "Batticaloa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ampara ---
ampara_data <- srilanka_weekly |> filter(district == "Ampara")
ampara_xts  <- xts(ampara_data$cases, order.by = ampara_data$start.date)
colnames(ampara_xts) <- "Cases"

g_ampara <- dygraph(ampara_xts, main = "Ampara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Trincomalee ---
trincomalee_data <- srilanka_weekly |> filter(district == "Trincomalee")
trincomalee_xts  <- xts(trincomalee_data$cases, order.by = trincomalee_data$start.date)
colnames(trincomalee_xts) <- "Cases"

g_trincomalee <- dygraph(trincomalee_xts, main = "Trincomalee: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kurunegala ---
kurunegala_data <- srilanka_weekly |> filter(district == "Kurunegala")
kurunegala_xts  <- xts(kurunegala_data$cases, order.by = kurunegala_data$start.date)
colnames(kurunegala_xts) <- "Cases"

g_kurunegala <- dygraph(kurunegala_xts, main = "Kurunegala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Puttalam ---
puttalam_data <- srilanka_weekly |> filter(district == "Puttalam")
puttalam_xts  <- xts(puttalam_data$cases, order.by = puttalam_data$start.date)
colnames(puttalam_xts) <- "Cases"

g_puttalam <- dygraph(puttalam_xts, main = "Puttalam: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Anuradhapura ---
anuradhapura_data <- srilanka_weekly |> filter(district == "Anuradhapura")
anuradhapura_xts  <- xts(anuradhapura_data$cases, order.by = anuradhapura_data$start.date)
colnames(anuradhapura_xts) <- "Cases"

g_anuradhapura <- dygraph(anuradhapura_xts, main = "Anuradhapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Polonnaruwa ---
polonnaruwa_data <- srilanka_weekly |> filter(district == "Polonnaruwa")
polonnaruwa_xts  <- xts(polonnaruwa_data$cases, order.by = polonnaruwa_data$start.date)
colnames(polonnaruwa_xts) <- "Cases"

g_polonnaruwa <- dygraph(polonnaruwa_xts, main = "Polonnaruwa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Badulla ---
badulla_data <- srilanka_weekly |> filter(district == "Badulla")
badulla_xts  <- xts(badulla_data$cases, order.by = badulla_data$start.date)
colnames(badulla_xts) <- "Cases"

g_badulla <- dygraph(badulla_xts, main = "Badulla: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Moneragala ---
moneragala_data <- srilanka_weekly |> filter(district == "Monaragala")
moneragala_xts  <- xts(moneragala_data$cases, order.by = moneragala_data$start.date)
colnames(moneragala_xts) <- "Cases"

g_moneragala <- dygraph(moneragala_xts, main = "Monaragala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ratnapura ---
ratnapura_data <- srilanka_weekly |> filter(district == "Ratnapura")
ratnapura_xts  <- xts(ratnapura_data$cases, order.by = ratnapura_data$start.date)
colnames(ratnapura_xts) <- "Cases"

g_ratnapura <- dygraph(ratnapura_xts, main = "Ratnapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kegalle ---
kegalle_data <- srilanka_weekly |> filter(district == "Kegalle")
kegalle_xts  <- xts(kegalle_data$cases, order.by = kegalle_data$start.date)
colnames(kegalle_xts) <- "Cases"

g_kegalle <- dygraph(kegalle_xts, main = "Kegalle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalmunai ---
kalmunai_data <- srilanka_weekly |> filter(district == "Kalmune")
kalmunai_xts  <- xts(kalmunai_data$cases, order.by = kalmunai_data$start.date)
colnames(kalmunai_xts) <- "Cases"

g_kalmunai <- dygraph(kalmunai_xts, main = "Kalmune: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")


plot_containers <- tags$div(
  style = "width: 100%; margin-top: 15px;",
  
  tags$div(id = "plot-colombo", class = "district-plot", style = "display: block;", g_colombo),
  tags$div(id = "plot-gampaha", class = "district-plot", style = "display: none;", g_gampaha),
  tags$div(id = "plot-kalutara", class = "district-plot", style = "display: none;", g_kalutara),
  tags$div(id = "plot-kandy", class = "district-plot", style = "display: none;", g_kandy),
  tags$div(id = "plot-matale", class = "district-plot", style = "display: none;", g_matale),
  tags$div(id = "plot-nuwara-eliya", class = "district-plot", style = "display: none;", g_nuwara_eliya),
  tags$div(id = "plot-galle", class = "district-plot", style = "display: none;", g_galle),
  tags$div(id = "plot-matara", class = "district-plot", style = "display: none;", g_matara),
  tags$div(id = "plot-hambantota", class = "district-plot", style = "display: none;", g_hambantota),
  tags$div(id = "plot-jaffna", class = "district-plot", style = "display: none;", g_jaffna),
  tags$div(id = "plot-kilinochchi", class = "district-plot", style = "display: none;", g_kilinochchi),
  tags$div(id = "plot-mannar", class = "district-plot", style = "display: none;", g_mannar),
  tags$div(id = "plot-vavuniya", class = "district-plot", style = "display: none;", g_vavuniya),
  tags$div(id = "plot-mullaitivu", class = "district-plot", style = "display: none;", g_mullaitivu),
  tags$div(id = "plot-batticaloa", class = "district-plot", style = "display: none;", g_batticaloa),
  tags$div(id = "plot-ampara", class = "district-plot", style = "display: none;", g_ampara),
  tags$div(id = "plot-trincomalee", class = "district-plot", style = "display: none;", g_trincomalee),
  tags$div(id = "plot-kurunegala", class = "district-plot", style = "display: none;", g_kurunegala),
  tags$div(id = "plot-puttalam", class = "district-plot", style = "display: none;", g_puttalam),
  tags$div(id = "plot-anuradhapura", class = "district-plot", style = "display: none;", g_anuradhapura),
  tags$div(id = "plot-polonnaruwa", class = "district-plot", style = "display: none;", g_polonnaruwa),
  tags$div(id = "plot-badulla", class = "district-plot", style = "display: none;", g_badulla),
  tags$div(id = "plot-moneragala", class = "district-plot", style = "display: none;", g_moneragala),
  tags$div(id = "plot-ratnapura", class = "district-plot", style = "display: none;", g_ratnapura),
  tags$div(id = "plot-kegalle", class = "district-plot", style = "display: none;", g_kegalle),
  tags$div(id = "plot-kalmunai", class = "district-plot", style = "display: none;", g_kalmunai)
)


dropdown_selector <- tags$select(
  id = "custom-district-selector",
  class = "form-select",
  style = "width: 100%; max-width: 300px; padding: 8px; border-radius: 5px; border: 1px solid #ccc; font-size: 14px;",
  onchange = "
    // Hide all plot containers
    document.querySelectorAll('.district-plot').forEach(el => el.style.display = 'none');
    
    // Show only the selected container
    const targetedPlot = document.getElementById(this.value);
    if (targetedPlot) { targetedPlot.style.display = 'block'; }
    
    // Force dygraphs to redraw perfectly inside the unhidden space
    window.dispatchEvent(new Event('resize'));
  ",
  tags$option(value = "plot-colombo", "Colombo"),
  tags$option(value = "plot-gampaha", "Gampaha"),
  tags$option(value = "plot-kalutara", "Kalutara"),
  tags$option(value = "plot-kandy", "Kandy"),
  tags$option(value = "plot-matale", "Matale"),
  tags$option(value = "plot-nuwara-eliya", "Nuwara Eliya"),
  tags$option(value = "plot-galle", "Galle"),
  tags$option(value = "plot-matara", "Matara"),
  tags$option(value = "plot-hambantota", "Hambanthota"),
  tags$option(value = "plot-jaffna", "Jaffna"),
  tags$option(value = "plot-kilinochchi", "Kilinochchi"),
  tags$option(value = "plot-mannar", "Mannar"),
  tags$option(value = "plot-vavuniya", "Vavuniya"),
  tags$option(value = "plot-mullaitivu", "Mullaitivu"),
  tags$option(value = "plot-batticaloa", "Batticaloa"),
  tags$option(value = "plot-ampara", "Ampara"),
  tags$option(value = "plot-trincomalee", "Trincomalee"),
  tags$option(value = "plot-kurunegala", "Kurunegala"),
  tags$option(value = "plot-puttalam", "Puttalam"),
  tags$option(value = "plot-anuradhapura", "Anuradhapura"),
  tags$option(value = "plot-polonnaruwa", "Polonnaruwa"),
  tags$option(value = "plot-badulla", "Badulla"),
  tags$option(value = "plot-moneragala", "Monaragala"),
  tags$option(value = "plot-ratnapura", "Ratnapura"),
  tags$option(value = "plot-kegalle", "Kegalle"),
  tags$option(value = "plot-kalmunai", "Kalmune")
)


browsable(
  tags$div(
    style = "padding: 10px; background: #fff; border-radius: 5px;",
    tags$label(tags$strong("Select District:")),
    tags$div(style = "margin-top: 5px;", dropdown_selector),
    plot_containers
  )
)
```

# District-wise Seasonal Plots

## Column {.tabset}

### District Overview

```{r}
#| label: Seasonal1
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20


srilanka_weekly_seasonal <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  ) |>
  filter(year_week >= 2020)


srilanka_weekly_seasonal <- srilanka_weekly_seasonal |>
  mutate(week_of_year = isoweek(week_start))  

g5 <- ggplot(srilanka_weekly_seasonal, 
                     aes(x = week_of_year, y = cases, group = year, color = factor(year))) +
  geom_line(size=2) +
  facet_wrap(vars(district), scales = "free_y", ncol = 5) +
  scale_x_continuous(breaks = seq(1, 52, by = 4)) + 
  labs(
    x = "Week of Year",
    y = "Weekly Dengue Cases",
    color = "Year",
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 26, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),
    legend.title = element_text(size = 26, face = "bold"),
    legend.text  = element_text(size = 24),
    legend.key.size = unit(1.5, "cm")
  ) + scale_color_brewer(palette = "Paired")

g5
```

### Individual District Analysis

```{r}
#| label: Seasonal2
#| echo: false
#| warning: false
#| message: false

seasonal <- function(dis_name) {

  dis <- srilanka_weekly_seasonal |>
    dplyr::filter(district == dis_name)

  ggplotly(
    ggplot(
      dis,
      aes(
        x = week_of_year,
        y = cases,
        group = year,
        color = factor(year)
      )
    ) +
      geom_line() +
      scale_x_continuous(breaks = seq(1, 52, by = 4)) +
      scale_color_viridis_d(option = "viridis") +
      labs(
        title = paste("Seasonal Plot -", dis_name),
        x = "Week of Year",
        y = "Weekly Dengue Cases",
        color = "Year"
      ) +
      theme_minimal()
  ) |> 
  layout(
    autosize = TRUE
  )
}
```

#### Column {.tabset}

##### Colombo

```{r}
seasonal("Colombo")
```

##### Gampaha

```{r}
seasonal("Gampaha")
```

##### Kalutara

```{r}
seasonal("Kalutara")
```

##### Kandy

```{r}
seasonal("Kandy")
```

##### Matale

```{r}
seasonal("Matale")
```

##### NuwaraEliya

```{r}
seasonal("NuwaraEliya")
```

##### Galle

```{r}
seasonal("Galle")
```

##### Hambanthota

```{r}
seasonal("Hambanthota")
```

##### Matara

```{r}
seasonal("Matara")
```

##### Jaffna

```{r}
seasonal("Jaffna")
```

##### Kilinochchi

```{r}
seasonal("Kilinochchi")
```

##### Mannar

```{r}
seasonal("Mannar")
```

##### Vavuniya

```{r}
seasonal("Vavuniya")
```

##### Mullaitivu

```{r}
seasonal("Mullaitivu")
```

##### Batticaloa

```{r}
seasonal("Batticaloa")
```

##### Ampara

```{r}
seasonal("Ampara")
```

##### Trincomalee

```{r}
seasonal("Trincomalee")
```

##### Kurunegala

```{r}
seasonal("Kurunegala")
```

##### Puttalam

```{r}
seasonal("Puttalam")
```

##### Anuradhapura

```{r}
seasonal("Anuradhapura")
```

##### Polonnaruwa

```{r}
seasonal("Polonnaruwa")
```

##### Badulla

```{r}
seasonal("Badulla")
```

##### Monaragala

```{r}
seasonal("Monaragala")
```

##### Ratnapura

```{r}
seasonal("Ratnapura")
```

##### Kegalle

```{r}
seasonal("Kegalle")
```

##### Kalmune

```{r}
seasonal("Kalmune")
```

# Dengue Patients Distribution

## Column {.tabset}

### Total Dengue Patients Distribution in 2025 (Last Year)

```{r}
#| label: Dis1
#| echo: false
#| warning: false
#| message: false

srilanka_2025 <- srilanka_weekly_data |>
  filter(year==2025) |> group_by(district) |>
  summarise(cases = sum(cases))

srilanka_2025 <- srilanka_2025 |>
  mutate(parents = "2025")

plot_ly(
    data = srilanka_2025,
     type="treemap",
     labels = ~ district,
     values = ~ cases,
    parents = ~ 2025,
    domain = list(column = 0),
    textinfo = "label+value+ percent parent"
  ) |> 
  layout(
    autosize = TRUE
  )

```

### Dengue Patients Distribution by Year and District

```{r}
#| label: Dis2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 80
#| fig-height: 30

ann_df <- srilanka_weekly_data |>
  group_by(year, district) |>        
  summarise(annual_cases = sum(cases), .groups = "drop")  
data(district)
ann_df <- ann_df |>
  mutate(district_clean = toupper(trimws(district)))

dis_sf <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

ann_df <- ann_df |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf$DISTRICT_clean[
      amatch(district_clean, district_sf$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()
ann_sf <- dis_sf |>
  left_join(ann_df, by = c("DISTRICT_clean" = "DISTRICT_match"))
ann_sf <- ann_sf |> filter(year > 2006)
ann_sf <- ann_sf |>
  mutate(
    cases_cat = cut(
      annual_cases,
      breaks = c(0, 30, 60, 100, 150, 250, 500, 1000, Inf),
      labels = c("0-30","31-60", "61-100", "101-150", "151-250",
                 "251-500", "501-1000", "Above 1000"),
      include.lowest = TRUE,
      right = TRUE
    )
  )


g6 <- ggplot(ann_sf) +
  geom_sf(aes(fill = cases_cat), color = "grey50", lwd = 0.2) +
  scale_fill_viridis_d(    
    option = "magma",
    direction = -1,
    na.value = "white",
    name = "Annual Dengue Cases"
  ) +
  facet_wrap(~year, ncol=10
             ) +
  theme_minimal() +
  coord_sf() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 26),
    strip.text = element_text(size = 28, face = "bold"),
    legend.title = element_text(size = 30, face = "bold"),
    legend.text  = element_text(size = 28),
    legend.key.size = unit(1.5, "cm"),
    plot.margin = margin(0, 0, 0, 0)
  )

g6

```

### Distribution of Yearly Dengue Patients for last 20 Years

```{r}
#| label: Dis3
#| echo: false
#| warning: false
#| message: false

ann_df <- ann_df |>
  filter(year > 2006) |>
  mutate(
    district = factor(district, levels = unique(district)),  
    cases_cat = cut(
      annual_cases,
      breaks = c(0, 30, 60, 100, 150, 250, 500, 1000, Inf),
      labels = c("0-30","31-60", "61-100", "101-150", "151-250",
                 "251-500", "501-1000", "Above 1000"),
      include.lowest = TRUE
    )
  )
g7 <- ggplot(ann_df) +
  geom_tile(aes(
    x = year, 
    y = district, 
    fill = cases_cat,
    text = paste0("District: ", district,
                  "<br>Year: ", year,
                  "<br>Annual Cases: ", annual_cases)
  ), color = "white") +
  scale_fill_viridis_d(
    option = "magma",  
    direction = -1,
    na.value = "grey90",
    name = "Annual Dengue Cases"
  ) +
  theme_minimal() +
  scale_x_continuous(
    breaks = unique(ann_df$year)  
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8)
  )
ggplotly(g7, tooltip = "text") |> 
  layout(
    autosize = TRUE
  )

```

# Climate-Dengue Relationship Analysis

## Column {.tabset}

### Time Series Analysis

```{r}
#| label: corr_ts
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20

srilanka_weekly_dengue <- srilanka_weekly_data |>
  mutate(
    start.date = mdy(start.date),
    end.date = mdy(end.date)
  )

srilanka_weekly_dengue <- srilanka_weekly_dengue |>
  mutate(
    Week_start = floor_date(start.date, unit = "week", week_start = 1)  # Monday as start of week
  ) 


srilanka_weekly_dengue <- srilanka_weekly_dengue |>
  mutate(
    year = year(Week_start),
    week = isoweek(Week_start)
  )


srilanka_weekly_dengue <- srilanka_weekly_dengue |>
  group_by(year, week) |>
  summarise(
    Cases = sum(cases, na.rm = TRUE),
    .groups = "drop"
  )


weekly_climate <- weekly_climate |>
  mutate(
    Year = year(Start_Date)
  )


climate_dengue_data <- weekly_climate |>
  left_join(
    srilanka_weekly_dengue,
    by = c("Year" = "year", "Week" = "week")
  )


climate_dengue_data <- climate_dengue_data |>
  select(
    Year,
    Week,
    Start_Date,
    Temperature,
    Rainfall,
    Relative_Humidity,
    Cases
  )

ts_data <- climate_dengue_data |>
  select(Start_Date, Temperature, Rainfall, Relative_Humidity, Cases) |>
  pivot_longer(
    cols = -Start_Date,
    names_to = "Variable",
    values_to = "Value"
  ) |>
  mutate(
    Variable = factor(Variable,
                      levels = c("Temperature",
                                 "Rainfall",
                                 "Relative_Humidity",
                                 "Cases"),
     labels = c("Temperature",
                "Rainfall",
                "Humidity (%)",
                "Cases"))
  )


g22 <- ggplot(ts_data, aes(x = Start_Date, y = Value, color = Variable)) +
  geom_line(linewidth = 0.5) +
  facet_grid(Variable ~ ., scales = "free_y") +
  scale_color_manual(values = c(
    "Temperature" = "steelblue",
    "Rainfall" = "forestgreen",
    "Humidity (%)" = "orange",
    "Cases" = "red"
  )) +
  labs(
    title = "Time Series of Climate Variables and Dengue Cases",
    x = "Time (Week Start Date)",
    y = "Value",
  ) +
  theme_minimal(base_size = 12) 

ggplotly(g22) |> 
  layout(
    autosize = TRUE
  )
```

### Scatterplot Analysis

#### Column {width="40%"}

```{r}
#| label: scatter1
#| echo: false
#| warning: false
#| message: false

g13 <- ggplot(climate_dengue_data,
       aes(x = Temperature, y = Cases)) +
  geom_point(color = "steelblue", alpha = 0.7, size = 2) +
  labs(
    title = "Temperature vs Weekly Dengue Cases",
    x = "Average Weekly Temperature (°C)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  )

ggplotly(g13)|> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: scatter2
#| echo: false
#| warning: false
#| message: false

g14 <- ggplot(climate_dengue_data,
       aes(x = Rainfall, y = Cases)) +
  geom_point(color = "forestgreen", alpha = 0.7, size = 2) +
  labs(
    title = "Weekly Rainfall vs Weekly Dengue Cases",
    x = "Total Weekly Rainfall (mm)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g14) |> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: scatter3
#| echo: false
#| warning: false
#| message: false

g15 <- ggplot(climate_dengue_data,
       aes(x = Relative_Humidity, y = Cases)) +
  geom_point(color = "darkorange", alpha = 0.7, size = 2) +
  labs(
    title = "Relative Humidity vs Weekly Dengue Cases",
    x = "Average Weekly Relative Humidity (%)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g15) |> 
  layout(
    autosize = TRUE
  )
```

### Lag 2 weeks scatterplots

#### Column {width="40%"}

```{r}
#| label: lag2_1
#| echo: false
#| warning: false
#| message: false

climate_dengue_data1 <- climate_dengue_data |>
  arrange(Year, Week)

climate_dengue_data1 <- climate_dengue_data1 %>%
  mutate(
    Temp_lag2 = lag(Temperature, 2),
    Temp_lag4 = lag(Temperature, 4),

    Rain_lag2 = lag(Rainfall, 2),
    Rain_lag4 = lag(Rainfall, 4),

    RH_lag2 = lag(Relative_Humidity, 2),
    RH_lag4 = lag(Relative_Humidity, 4)
  )

climate_dengue_data_lag <- climate_dengue_data1 %>%
  filter(
    !is.na(Rain_lag4),
    !is.na(Temp_lag4),
    !is.na(RH_lag4)
  )

g16 <- ggplot(climate_dengue_data_lag, aes(x = Temp_lag2, y = Cases)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  labs(
    title = "Lag 2 Weeks: Temperature vs Dengue Cases",
    x = "Average Temperature (Lag 2 Weeks, °C)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g16) |> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: lag2_2
#| echo: false
#| warning: false
#| message: false

g17 <- ggplot(climate_dengue_data_lag, aes(x = Rain_lag2, y = Cases)) +
  geom_point(alpha = 0.7, color = "forestgreen") +
  labs(
    title = "Lag 2 Weeks: Rainfall vs Dengue Cases",
    x = "Total Rainfall (Lag 2 Weeks, mm)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g17) |> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: lag2_3
#| echo: false
#| warning: false
#| message: false

g18 <- ggplot(climate_dengue_data_lag, aes(x = RH_lag2, y = Cases)) +
  geom_point(alpha = 0.7, color = "orange") +
  labs(
    title = "Lag 2 Weeks: Relative Humidity vs Dengue Cases",
    x = "Relative Humidity (Lag 2 Weeks, %)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g18) |> 
  layout(
    autosize = TRUE
  )
```

### Lag 4 weeks scatterplots

#### Column {width="40%"}

```{r}
#| label: lag4_1
#| echo: false
#| warning: false
#| message: false

g19 <- ggplot(climate_dengue_data_lag, aes(x = Temp_lag4, y = Cases)) +
  geom_point(alpha = 0.7, color = "darkblue") +
  labs(
    title = "Lag 4 Weeks: Temperature vs Dengue Cases",
    x = "Average Temperature (Lag 4 Weeks, °C)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g19) |> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: lag4_2
#| echo: false
#| warning: false
#| message: false

g20 <- ggplot(climate_dengue_data_lag, aes(x = Rain_lag4, y = Cases)) +
  geom_point(alpha = 0.7, color = "darkgreen") +
  labs(
    title = "Lag 4 Weeks: Rainfall vs Dengue Cases",
    x = "Total Rainfall (Lag 4 Weeks, mm)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g20) |> 
  layout(
    autosize = TRUE
  )
```

#### Column {width="40%"}

```{r}
#| label: lag4_3
#| echo: false
#| warning: false
#| message: false

g21 <- ggplot(climate_dengue_data_lag, aes(x = RH_lag4, y = Cases)) +
  geom_point(alpha = 0.7, color = "darkorange") +
  labs(
    title = "Lag 4 Weeks: Relative Humidity vs Dengue Cases",
    x = "Relative Humidity (Lag 4 Weeks, %)",
    y = "Weekly Dengue Cases"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 10,
      hjust = 0.5
    )
  ) 

ggplotly(g21) |> 
  layout(
    autosize = TRUE
  )
```

### Detrended Cross-Correlation Analysis (DCCA)

```{r}
#| label: DCCA
#| echo: false
#| warning: false
#| message: false


temp <- climate_dengue_data$Temperature
rain <- climate_dengue_data$Rainfall
humidity <- climate_dengue_data$Relative_Humidity
dengue <- climate_dengue_data$Cases


calculate_dcca <- function(x, y, scales){

  df <- data.frame(x = x, y = y)

  df <- df[
    complete.cases(df) &
      is.finite(df$x) &
      is.finite(df$y),
  ]

  x <- df$x
  y <- df$y

  rho_values <- numeric(length(scales))

  for(i in seq_along(scales)){

    rho_values[i] <- rhodcca(
      y1 = x,
      y2 = y,
      m = scales[i],
      nu = 0,
      overlap = TRUE
    )$rhodcca

  }

  data.frame(
    Scale = scales,
    rho_DCCA = rho_values
  )
}



scales <- seq(4, 52, by = 2)


dcca_temp <- calculate_dcca(
  temp,
  dengue,
  scales
)

dcca_temp$Variable <- "Temperature"



dcca_rain <- calculate_dcca(
  rain,
  dengue,
  scales
)

dcca_rain$Variable <- "Rainfall"



dcca_hum <- calculate_dcca(
  humidity,
  dengue,
  scales
)

dcca_hum$Variable <- "Humidity"



dcca_results <- rbind(
  dcca_temp,
  dcca_rain,
  dcca_hum
)



g23 <- ggplot(
  dcca_results,
  aes(
    x = Scale,
    y = rho_DCCA,
    color = Variable
  )
) +

geom_point(size = 2) +

geom_line(linewidth = 0.8) +

geom_hline(
  yintercept = 0,
  linetype = "dashed",
  linewidth = 0.6
) +

scale_color_manual(
  values = c(
    "Temperature" = "steelblue",
    "Rainfall" = "forestgreen",
    "Humidity" = "orange"
  )
) +

scale_x_continuous(
  breaks = seq(4,52,by=4)
) +

labs(
  title = "Detrended Cross-Correlation between Climate Variables and Dengue Cases",
  subtitle = "Scale-dependent ρDCCA analysis (4–52 weeks)",
  x = "Time Scale (Weeks)",
  y = expression(rho[DCCA]),
  color = "Climate Variable"
) +

theme_minimal(base_size = 12) +

theme(
  legend.position = "bottom"
)


ggplotly(g23) |> 
  layout(
    autosize = TRUE
  )

```

# Global Comparison

## Column {.tabset}

### Global Distribution of the Patient in 2025

```{r}
#| label: global1
#| echo: false
#| warning: false
#| message: false

Global <- Global |>
  mutate(year = year(date))

global_2025 <- Global |>
  filter(year == 2025) |>
  group_by(country) |>
  summarise(cases = sum(cases), deaths = sum(deaths)) |>
  arrange(desc(cases))

world <- ne_countries(scale = "medium", returnclass = "sf")

world2 <- world |>
  left_join(
    global_2025,
    by = c("name" = "country")
  )

world2 <- world2 |>
  mutate(
    cases_cat = cut(
      cases,
      breaks = c(-1, 0, 499, 4999, 49999, Inf),
      labels = c(
        "0",
        "1-499",
        "500-4,999",
        "5,000-49,999",
        "50,000+"
      ),
      include.lowest = TRUE
    )
  )

pal <- colorFactor(
  palette = c(
    "#F6EDB3",
    "#FDC777",
    "#F89C59",
    "#F15B4A",
    "#C91F5C"
  ),
  domain = world2$cases_cat,
  na.color = "#EEEEEE"
)

g8 <- leaflet(
  world2,
  options = leafletOptions(
    worldCopyJump = FALSE,
    minZoom = 2
  )) |>
  
  addProviderTiles(
    "CartoDB.Positron",
    options = providerTileOptions(noWrap = TRUE)
  )  |>

     addPolygons(
       fillColor = ~pal(cases_cat),
       color = "#cccccc",
       weight = 0.5,
       fillOpacity = 0.9,

    popup = ~paste0(
      "<b>", name, "</b><br>",
      "Cases: ", cases
    )
  ) |>

  addLegend(
    pal = pal,
    values = ~cases_cat,
    title = "Total Cases",
    position = "bottomright"
  )

g8 <- g8 |>
  htmlwidgets::onRender("
    function(el,x){
      this.invalidateSize();
    }
  ")

g8
```

### Sri Lanka vs. Top 10 Countries Worldwide (2025)

#### Column {width="50%"}

```{r}
#| label: global2
#| echo: false
#| warning: false
#| message: false

global_top10 <- global_2025 |>
  arrange(desc(cases)) |>
  slice_head(n = 10)

sri_lanka <- global_2025 |>
  filter(country == "Sri Lanka")

global_top10 <- bind_rows(global_top10, sri_lanka) |>
  distinct(country, .keep_all = TRUE)

global_top10$country <- factor(
  global_top10$country,
  levels = rev(global_top10$country)
)

global_top10 <- global_top10 |> 
  pivot_longer( cols = c(cases, deaths), names_to = "type", values_to = "count" ) |>
  mutate(
    count = replace_na(count, 0)
  )


g9 <- plot_ly(
  global_top10,
  x = ~count,
  y = ~country,
  type = "bar",
  color = ~type,
  colors = c(
    "cases" = "steelblue",
    "deaths" = "firebrick"
  )
) |>
  layout(
    title = list(
      text = "Total Dengue Cases and Deaths by Country (2025)",
      font = list(size = 16)
    ),
    xaxis = list(title = "Count"),
    yaxis = list(title = list(text = "Country", standoff = 10)),
    barmode = "stack",
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5
    )
  )

g9
```

#### Column {width="50%"}

```{r}
#| label: global3
#| echo: false
#| warning: false
#| message: false


global_top10_rate <- global_2025 |>
  arrange(desc(cases)) |>
  slice_head(n = 10)

sri_lanka <- global_2025 |>
  filter(country == "Sri Lanka")

global_top10_rate <- bind_rows(global_top10_rate, sri_lanka) |>
  distinct(country, .keep_all = TRUE)

global_top10_rate <- global_top10_rate |>
  mutate(
    deaths = replace_na(deaths, 0)
  ) |>
  mutate(
    rate = (deaths/cases) * 100
  )

global_top10_rate$country <- factor(
  global_top10_rate$country,
  levels = rev(global_top10_rate$country)
)

g10 <- plot_ly(
  global_top10_rate,
  x = ~rate,
  y = ~reorder(country, rate),
  type = "bar",
  marker = list(color = "#8B0000"),
  name = "Death Rate",
  showlegend = FALSE
) |>
  layout(
    title = list(text = "Dengue Death Rate by Country (2025)", size = 16),
    xaxis = list(title = 'Death Rate (%)'),
    yaxis = list(
     title = list(text = "Country", standoff = 10)
    )
  )

g10 |> 
  layout(
    autosize = TRUE
  )
```

### Sri Lanka vs. South Eastern Asian Countries-WHO Region (2025)

#### Column {width="50%"}

```{r}
#| label: global4
#| echo: false
#| warning: false
#| message: false

global_2025_asia <- Global |>
  filter(year == 2025 & who_region == "SEAR") |>
  group_by(country) |>
  summarise(cases = sum(cases), deaths = sum(deaths)) |>
  arrange(desc(cases))

global_2025_asia$country <- factor(
  global_2025_asia$country,
  levels = rev(global_2025_asia$country)
)

global_2025_asia_new <- global_2025_asia |> 
  pivot_longer( cols = c(cases, deaths), names_to = "type", values_to = "count" ) |>
  mutate(
    count = replace_na(count, 0)
  )

g11 <- plot_ly(
  global_2025_asia_new,
  x = ~count,
  y = ~country,
  type = "bar",
  color = ~type,
  colors = c(
    "cases" = "steelblue",
    "deaths" = "firebrick"
  )
) |>
  layout(
    title = list(
      text = "Total Dengue Cases and Deaths by Country (2025)",
      font = list(size = 16)
    ),
    xaxis = list(title = "Count"),
    yaxis = list(title = list(text = "Country", standoff = 10)),
    barmode = "stack",
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5
    )
  )

g11 |> 
  layout(
    autosize = TRUE
  )

```

#### Column {width="50%"}

```{r}
#| label: global5
#| echo: false
#| warning: false
#| message: false

global_2025_asia_rate <- global_2025_asia |>
  mutate(
    deaths = replace_na(deaths, 0)
  ) |>
  mutate(
    rate = (deaths/cases) * 100
  )

global_2025_asia_rate$country <- factor(
  global_2025_asia_rate$country,
  levels = rev(global_2025_asia_rate$country)
)

g12 <- plot_ly(
  global_2025_asia_rate,
  x = ~rate,
  y = ~reorder(country, rate),
  type = "bar",
  marker = list(color = "#8B0000"),
  name = "Death Rate",
  showlegend = FALSE
) |>
  layout(
    title = list(text = "Dengue Death Rate by Country (2025)", size = 16),
    xaxis = list(title = 'Death Rate (%)'),
    yaxis = list(
     title = list(text = "Country", standoff = 10)
    )
  )

g12 |> 
  layout(
    autosize = TRUE
  )
```

# About

This dashboard provides an interactive overview of the dengue situation in Sri Lanka using the latest available surveillance data. Built with R using the Quarto Dashboard framework, it presents recent dengue trends, district-level analyses, historical patterns, comparisons with the global dengue situation, and explores the relationship between dengue transmission and key climate variables. The dashboard offers valuable insights for monitoring, understanding, and supporting evidence-based decision-making on dengue dynamics in Sri Lanka.

**Created by**

-   [Y.M. Amali P. Rajapaksha](https://amalipriyanwada.netlify.app/), Department of Statistics, University of Sri Jayewardenepura.

**Code**

-   The code behind the dashboard is available on [GitHub](https://github.com/amalirajapaksha/denguedashboard_SriLanka).

**Data**

-   All local data up to December 2025 used in this dashboard are obtained from the weekly epidemiological reports published by the Epidemiology Unit, Ministry of Health, Sri Lanka, through the `denguedatahub` R package, available at https://github.com/thiyangt/denguedatahub.

-   Data for 2026 are obtained from the weekly epidemiological reports published by the Epidemiology Unit, Ministry of Health, Sri Lanka, and the Weekly Dengue Updates published by the National Dengue Control Unit, Ministry of Health, Sri Lanka. These data are extracted through web scraping using the `convert_slwer_to_tidy()` function available in the `denguedatahub` package.

-   Sri Lanka maps are created using the `ceylon` package in R available here https://github.com/thiyangt/ceylon.

-   The global data used in this dashboard are obtained from the World Health Organization's [Global Dengue Surveillance Dashboard](https://worldhealthorg.shinyapps.io/dengue_global/).

-   The climate data used in this study were downloaded from the [NASA POWER Data Access Viewer](https://power.larc.nasa.gov/data-access-viewer/), provided by the National Aeronautics and Space Administration (NASA).

**Remarks**

-   Although the Ampara district has been divided into two Regional Directorates of Health Services (RDHS) areas, Ampara and Kalmunai, both Kalmunai and Ampara RDHS counts are combined and represented under Ampara district in the Sri Lanka country map.

-   Countries with the highest number of confirmed cases were selected as top 10 countries in Global Comparison tab.

**References**

-   **Visualization techniques:** Referenced the [Sri Lanka COVID-19 Dashboard](https://thiyangt.github.io/coviddashboard/#about) to explore and adapt visualization techniques that can be applied for disease surveillance and epidemiological analysis, including interactive summaries and temporal trend visualization.

-   **Climate–dengue correlation analysis methods:** Reviewed statistical approaches from the papers *"Analysis of the correlation between climatic variables and Dengue cases in the city of Alagoinhas/BA"* by Marcos Batista Figueredo et al. and *"Correlation between incidence of dengue and climatic factors in the Philippines: An ecological study"* by Ann Kashmer D. Yu et al. to identify suitable methods for analysing relationships between climatic variables and dengue incidence.
