---
title: "Sri Lanka Dengue Dashboard"
format:
  dashboard:
    css: styles.css
editor: source
---

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(denguedatahub)
library(ceylon)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(here)
library(stringdist)
library(sf)
library(viridis)
library(cartogram)
library(ISOweek)
library(lubridate)
library(xts)
library(dygraphs)
library(crosstalk)
library(leaflet)
library(htmltools)
library(plotly)
library(shiny)
library(RColorBrewer)
library(scales) # For percentage formatting
library(RColorBrewer) # For Dark2 color palette
library(GGally)
```

```{r}
#| label: load-data
#| message: false
#| warning: false
data(srilanka_weekly_data)

weekly_2026     <- read_csv(here("data","sri_lanka_weekly_dengue_cases_2026.csv"))
weekly_2026$year <- as.numeric(weekly_2026$year)
weekly_2026$week <- as.numeric(weekly_2026$week)
weekly_2026$start.date <- as.Date(weekly_2026$start.date)
weekly_2026$district <- as.character(weekly_2026$district)
weekly_2026$cases <- as.numeric(weekly_2026$cases)
```

```{r}
#| label: cleandata

weekly_cases_by_date <- weekly_2026 |>
  group_by(start.date) |>
  summarise(weekly_cases = sum(cases, na.rm = TRUE)) |>
  arrange(start.date)
```

# Overview

**Latest Dengue Surveillance Summary**\
**Last Updated:** `r format(Sys.Date(), "%A, %B %d, %Y")` \| **Latest local data available for:** May 17, 2026

## Row {height="20%"}

```{r}
#| content: valuebox
#| title: "Total Cases in 2026"
#| color: "#d95f02"
#| icon: "virus"

sum(weekly_2026$cases, na.rm = TRUE)
```

```{r}
#| content: valuebox
#| title: "Average Weekly Cases in 2026"
#| color: "#7570b3"
#| icon: "calendar-week"

round(sum(weekly_2026$cases, na.rm = TRUE)/20)
```

```{r}
#| content: valuebox
#| title: "Highest Risk District in 2026"
#| icon: "geo-alt"
#| color: "#1b9e77"

weekly_2026 |>
  group_by(district) |>
  summarise(total_cases = sum(cases, na.rm = TRUE)) |>
  arrange(desc(total_cases)) |>
  slice(1) |>
  pull(district)
```

## Row {height="65%"}

### Column {width="40%"}

```{r}
#| label: map1
#| echo: false
#| message: false
#| warning: false


district_df <- weekly_2026 |>
  group_by(district) |>       
  summarise(Count = sum(cases), .groups = "drop")  

data(district)
district_df <- district_df |>
  mutate(district_clean = toupper(trimws(district)))

district_sf <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df <- district_df |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf$DISTRICT_clean[
      amatch(district_clean, district_sf$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2 <- as.data.frame(district_df)

map_data <- left_join(
  district_sf,
  district_df2,
  by = c("DISTRICT_clean" = "DISTRICT_match")
)


g1 <- ggplotly(ggplot(map_data) +
  geom_sf(aes(fill = Count, text = paste('District:', district))) +
  labs(title = "Map for Total Dengue Cases in 2026") +  
  scale_fill_viridis() +
  theme_minimal()+
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(0, 0, 0, 0)
  ))

g1 <- g1 |>
  layout(
    title = list(
      text = "Map for Distribution of Total Dengue Cases in 2026",
      font = list(size = 16)   
    ))

g1

```

### Column {width="40%"}

```{r}
#| label: map2
#| echo: false
#| message: false
#| warning: false


map_data <- map_data |>
mutate(
risk = Count / population
)

map_data <- map_data |>
filter(!is.na(risk) & risk > 0)

dorl <- cartogram_dorling(map_data, weight = "population") |>
  st_as_sf()

g2 <- ggplotly(
  ggplot(dorl) +
    geom_sf(data = map_data) +
    geom_sf(aes(
      fill = Count,
      text = paste(
        "District:", DISTRICT_clean
      )
    )) +
    labs(title = "Map for Dengue Cases in 2026 weighted by population") +
    scale_fill_viridis()+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
)

g2 <- g2 |>
    layout(
    title = list(
       text = "Map for Dengue Cases in 2026 weighted by population",
       font = list(size = 15)   
     ))

g2
```

### Column {width="40%"}

```{r}
#| label: TS1
#| echo: false
#| message: false
#| warning: false

g3 <- ggplotly(ggplot(weekly_cases_by_date, aes(x = start.date, y = weekly_cases,)) +
      geom_line(size = 1, col = "#3b0f70") +           
      geom_point(size = 1, col = "#3b0f70") +
      scale_color_viridis_d() +
      scale_x_date(
    date_labels = "%b %d",   
    date_breaks = "2 weeks") +
      labs(
        title = "Weekly Dengue Cases in 2026",
        x = "Date (Month-Day)",
        y = "Cases"
      ) +
      theme_minimal(base_size = 14)
    )

g3 <- g3 |>
  layout(
    title = list(
      text = "Weekly Dengue Cases in 2026",
      font = list(size = 16)   
    ),

    xaxis = list(
      title = "Date (Month-Day)",
      titlefont = list(size = 14),   
      tickfont = list(size = 10),    
      tickangle = -60                  
    ),

    yaxis = list(
      title = "Cases",
      titlefont = list(size = 14),
      tickfont = list(size = 12),
      range = c(0, 2000)
    )
  )

g3
```

# District-wise Time Series Trends

## Column {.tabset}

### Overview of District Trends


```{r}
#| label: TS2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20

 
srilanka_weekly_df <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  )

g4 <- ggplot(srilanka_weekly_df, aes(x = week_start, y = cases)) +
  geom_line(color = "darkblue", linewidth = 0.8) +
  facet_wrap(vars(district), scales = "free_y", ncol = 4) + 
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    x = "Year",
    y = "Weekly dengue cases"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 26, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28)
  )

g4
```


### Explore Individual District Trends

```{r}
#| label: TS3
#| echo: false
#| warning: false
#| message: false


srilanka_weekly <- srilanka_weekly_data |>
  mutate(start.date = mdy(start.date)) |>
  arrange(start.date)

# --- Colombo ---
colombo_data <- srilanka_weekly |> filter(district == "Colombo")
colombo_xts  <- xts(colombo_data$cases, order.by = colombo_data$start.date)
colnames(colombo_xts) <- "Cases"

g_colombo <- dygraph(colombo_xts, main = "Colombo: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Gampaha ---
gampaha_data <- srilanka_weekly |> filter(district == "Gampaha")
gampaha_xts  <- xts(gampaha_data$cases, order.by = gampaha_data$start.date)
colnames(gampaha_xts) <- "Cases"

g_gampaha <- dygraph(gampaha_xts, main = "Gampaha: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalutara ---
kalutara_data <- srilanka_weekly |> filter(district == "Kalutara")
kalutara_xts  <- xts(kalutara_data$cases, order.by = kalutara_data$start.date)
colnames(kalutara_xts) <- "Cases"

g_kalutara <- dygraph(kalutara_xts, main = "Kalutara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kandy ---
kandy_data <- srilanka_weekly |> filter(district == "Kandy")
kandy_xts  <- xts(kandy_data$cases, order.by = kandy_data$start.date)
colnames(kandy_xts) <- "Cases"

g_kandy <- dygraph(kandy_xts, main = "Kandy: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matale ---
matale_data <- srilanka_weekly |> filter(district == "Matale")
matale_xts  <- xts(matale_data$cases, order.by = matale_data$start.date)
colnames(matale_xts) <- "Cases"

g_matale <- dygraph(matale_xts, main = "Matale: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Nuwara Eliya ---
nuwara_eliya_data <- srilanka_weekly |> filter(district == "NuwaraEliya")
nuwara_eliya_xts  <- xts(nuwara_eliya_data$cases, order.by = nuwara_eliya_data$start.date)
colnames(nuwara_eliya_xts) <- "Cases"

g_nuwara_eliya <- dygraph(nuwara_eliya_xts, main = "Nuwara Eliya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Galle ---
galle_data <- srilanka_weekly |> filter(district == "Galle")
galle_xts  <- xts(galle_data$cases, order.by = galle_data$start.date)
colnames(galle_xts) <- "Cases"

g_galle <- dygraph(galle_xts, main = "Galle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matara ---
matara_data <- srilanka_weekly |> filter(district == "Matara")
matara_xts  <- xts(matara_data$cases, order.by = matara_data$start.date)
colnames(matara_xts) <- "Cases"

g_matara <- dygraph(matara_xts, main = "Matara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Hambantota ---
hambantota_data <- srilanka_weekly |> filter(district == "Hambanthota")
hambantota_xts  <- xts(hambantota_data$cases, order.by = hambantota_data$start.date)
colnames(hambantota_xts) <- "Cases"

g_hambantota <- dygraph(hambantota_xts, main = "Hambanthota: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Jaffna ---
jaffna_data <- srilanka_weekly |> filter(district == "Jaffna")
jaffna_xts  <- xts(jaffna_data$cases, order.by = jaffna_data$start.date)
colnames(jaffna_xts) <- "Cases"

g_jaffna <- dygraph(jaffna_xts, main = "Jaffna: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kilinochchi ---
kilinochchi_data <- srilanka_weekly |> filter(district == "Kilinochchi")
kilinochchi_xts  <- xts(kilinochchi_data$cases, order.by = kilinochchi_data$start.date)
colnames(kilinochchi_xts) <- "Cases"

g_kilinochchi <- dygraph(kilinochchi_xts, main = "Kilinochchi: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mannar ---
mannar_data <- srilanka_weekly |> filter(district == "Mannar")
mannar_xts  <- xts(mannar_data$cases, order.by = mannar_data$start.date)
colnames(mannar_xts) <- "Cases"

g_mannar <- dygraph(mannar_xts, main = "Mannar: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Vavuniya ---
vavuniya_data <- srilanka_weekly |> filter(district == "Vavuniya")
vavuniya_xts  <- xts(vavuniya_data$cases, order.by = vavuniya_data$start.date)
colnames(vavuniya_xts) <- "Cases"

g_vavuniya <- dygraph(vavuniya_xts, main = "Vavuniya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mullaitivu ---
mullaitivu_data <- srilanka_weekly |> filter(district == "Mullaitivu")
mullaitivu_xts  <- xts(mullaitivu_data$cases, order.by = mullaitivu_data$start.date)
colnames(mullaitivu_xts) <- "Cases"

g_mullaitivu <- dygraph(mullaitivu_xts, main = "Mullaitivu: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Batticaloa ---
batticaloa_data <- srilanka_weekly |> filter(district == "Batticaloa")
batticaloa_xts  <- xts(batticaloa_data$cases, order.by = batticaloa_data$start.date)
colnames(batticaloa_xts) <- "Cases"

g_batticaloa <- dygraph(batticaloa_xts, main = "Batticaloa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ampara ---
ampara_data <- srilanka_weekly |> filter(district == "Ampara")
ampara_xts  <- xts(ampara_data$cases, order.by = ampara_data$start.date)
colnames(ampara_xts) <- "Cases"

g_ampara <- dygraph(ampara_xts, main = "Ampara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Trincomalee ---
trincomalee_data <- srilanka_weekly |> filter(district == "Trincomalee")
trincomalee_xts  <- xts(trincomalee_data$cases, order.by = trincomalee_data$start.date)
colnames(trincomalee_xts) <- "Cases"

g_trincomalee <- dygraph(trincomalee_xts, main = "Trincomalee: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kurunegala ---
kurunegala_data <- srilanka_weekly |> filter(district == "Kurunegala")
kurunegala_xts  <- xts(kurunegala_data$cases, order.by = kurunegala_data$start.date)
colnames(kurunegala_xts) <- "Cases"

g_kurunegala <- dygraph(kurunegala_xts, main = "Kurunegala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Puttalam ---
puttalam_data <- srilanka_weekly |> filter(district == "Puttalam")
puttalam_xts  <- xts(puttalam_data$cases, order.by = puttalam_data$start.date)
colnames(puttalam_xts) <- "Cases"

g_puttalam <- dygraph(puttalam_xts, main = "Puttalam: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Anuradhapura ---
anuradhapura_data <- srilanka_weekly |> filter(district == "Anuradhapura")
anuradhapura_xts  <- xts(anuradhapura_data$cases, order.by = anuradhapura_data$start.date)
colnames(anuradhapura_xts) <- "Cases"

g_anuradhapura <- dygraph(anuradhapura_xts, main = "Anuradhapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Polonnaruwa ---
polonnaruwa_data <- srilanka_weekly |> filter(district == "Polonnaruwa")
polonnaruwa_xts  <- xts(polonnaruwa_data$cases, order.by = polonnaruwa_data$start.date)
colnames(polonnaruwa_xts) <- "Cases"

g_polonnaruwa <- dygraph(polonnaruwa_xts, main = "Polonnaruwa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Badulla ---
badulla_data <- srilanka_weekly |> filter(district == "Badulla")
badulla_xts  <- xts(badulla_data$cases, order.by = badulla_data$start.date)
colnames(badulla_xts) <- "Cases"

g_badulla <- dygraph(badulla_xts, main = "Badulla: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Moneragala ---
moneragala_data <- srilanka_weekly |> filter(district == "Monaragala")
moneragala_xts  <- xts(moneragala_data$cases, order.by = moneragala_data$start.date)
colnames(moneragala_xts) <- "Cases"

g_moneragala <- dygraph(moneragala_xts, main = "Monaragala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ratnapura ---
ratnapura_data <- srilanka_weekly |> filter(district == "Ratnapura")
ratnapura_xts  <- xts(ratnapura_data$cases, order.by = ratnapura_data$start.date)
colnames(ratnapura_xts) <- "Cases"

g_ratnapura <- dygraph(ratnapura_xts, main = "Ratnapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kegalle ---
kegalle_data <- srilanka_weekly |> filter(district == "Kegalle")
kegalle_xts  <- xts(kegalle_data$cases, order.by = kegalle_data$start.date)
colnames(kegalle_xts) <- "Cases"

g_kegalle <- dygraph(kegalle_xts, main = "Kegalle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalmunai ---
kalmunai_data <- srilanka_weekly |> filter(district == "Kalmune")
kalmunai_xts  <- xts(kalmunai_data$cases, order.by = kalmunai_data$start.date)
colnames(kalmunai_xts) <- "Cases"

g_kalmunai <- dygraph(kalmunai_xts, main = "Kalmune: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")


plot_containers <- tags$div(
  style = "width: 100%; margin-top: 15px;",
  
  tags$div(id = "plot-colombo", class = "district-plot", style = "display: block;", g_colombo),
  tags$div(id = "plot-gampaha", class = "district-plot", style = "display: none;", g_gampaha),
  tags$div(id = "plot-kalutara", class = "district-plot", style = "display: none;", g_kalutara),
  tags$div(id = "plot-kandy", class = "district-plot", style = "display: none;", g_kandy),
  tags$div(id = "plot-matale", class = "district-plot", style = "display: none;", g_matale),
  tags$div(id = "plot-nuwara-eliya", class = "district-plot", style = "display: none;", g_nuwara_eliya),
  tags$div(id = "plot-galle", class = "district-plot", style = "display: none;", g_galle),
  tags$div(id = "plot-matara", class = "district-plot", style = "display: none;", g_matara),
  tags$div(id = "plot-hambantota", class = "district-plot", style = "display: none;", g_hambantota),
  tags$div(id = "plot-jaffna", class = "district-plot", style = "display: none;", g_jaffna),
  tags$div(id = "plot-kilinochchi", class = "district-plot", style = "display: none;", g_kilinochchi),
  tags$div(id = "plot-mannar", class = "district-plot", style = "display: none;", g_mannar),
  tags$div(id = "plot-vavuniya", class = "district-plot", style = "display: none;", g_vavuniya),
  tags$div(id = "plot-mullaitivu", class = "district-plot", style = "display: none;", g_mullaitivu),
  tags$div(id = "plot-batticaloa", class = "district-plot", style = "display: none;", g_batticaloa),
  tags$div(id = "plot-ampara", class = "district-plot", style = "display: none;", g_ampara),
  tags$div(id = "plot-trincomalee", class = "district-plot", style = "display: none;", g_trincomalee),
  tags$div(id = "plot-kurunegala", class = "district-plot", style = "display: none;", g_kurunegala),
  tags$div(id = "plot-puttalam", class = "district-plot", style = "display: none;", g_puttalam),
  tags$div(id = "plot-anuradhapura", class = "district-plot", style = "display: none;", g_anuradhapura),
  tags$div(id = "plot-polonnaruwa", class = "district-plot", style = "display: none;", g_polonnaruwa),
  tags$div(id = "plot-badulla", class = "district-plot", style = "display: none;", g_badulla),
  tags$div(id = "plot-moneragala", class = "district-plot", style = "display: none;", g_moneragala),
  tags$div(id = "plot-ratnapura", class = "district-plot", style = "display: none;", g_ratnapura),
  tags$div(id = "plot-kegalle", class = "district-plot", style = "display: none;", g_kegalle),
  tags$div(id = "plot-kalmunai", class = "district-plot", style = "display: none;", g_kalmunai)
)


dropdown_selector <- tags$select(
  id = "custom-district-selector",
  class = "form-select",
  style = "width: 100%; max-width: 300px; padding: 8px; border-radius: 5px; border: 1px solid #ccc; font-size: 14px;",
  onchange = "
    // Hide all plot containers
    document.querySelectorAll('.district-plot').forEach(el => el.style.display = 'none');
    
    // Show only the selected container
    const targetedPlot = document.getElementById(this.value);
    if (targetedPlot) { targetedPlot.style.display = 'block'; }
    
    // Force dygraphs to redraw perfectly inside the unhidden space
    window.dispatchEvent(new Event('resize'));
  ",
  tags$option(value = "plot-colombo", "Colombo"),
  tags$option(value = "plot-gampaha", "Gampaha"),
  tags$option(value = "plot-kalutara", "Kalutara"),
  tags$option(value = "plot-kandy", "Kandy"),
  tags$option(value = "plot-matale", "Matale"),
  tags$option(value = "plot-nuwara-eliya", "Nuwara Eliya"),
  tags$option(value = "plot-galle", "Galle"),
  tags$option(value = "plot-matara", "Matara"),
  tags$option(value = "plot-hambantota", "Hambanthota"),
  tags$option(value = "plot-jaffna", "Jaffna"),
  tags$option(value = "plot-kilinochchi", "Kilinochchi"),
  tags$option(value = "plot-mannar", "Mannar"),
  tags$option(value = "plot-vavuniya", "Vavuniya"),
  tags$option(value = "plot-mullaitivu", "Mullaitivu"),
  tags$option(value = "plot-batticaloa", "Batticaloa"),
  tags$option(value = "plot-ampara", "Ampara"),
  tags$option(value = "plot-trincomalee", "Trincomalee"),
  tags$option(value = "plot-kurunegala", "Kurunegala"),
  tags$option(value = "plot-puttalam", "Puttalam"),
  tags$option(value = "plot-anuradhapura", "Anuradhapura"),
  tags$option(value = "plot-polonnaruwa", "Polonnaruwa"),
  tags$option(value = "plot-badulla", "Badulla"),
  tags$option(value = "plot-moneragala", "Monaragala"),
  tags$option(value = "plot-ratnapura", "Ratnapura"),
  tags$option(value = "plot-kegalle", "Kegalle"),
  tags$option(value = "plot-kalmunai", "Kalmune")
)


browsable(
  tags$div(
    style = "padding: 10px; background: #fff; border-radius: 5px;",
    tags$label(tags$strong("Select District:")),
    tags$div(style = "margin-top: 5px;", dropdown_selector),
    plot_containers
  )
)
```

# District-wise Seasonal Plots

## Column {.tabset}

### District Overview

```{r}
#| label: Seasonal1
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20


srilanka_weekly_seasonal <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  ) |>
  filter(year_week >= 2020)


srilanka_weekly_seasonal <- srilanka_weekly_seasonal |>
  mutate(week_of_year = isoweek(week_start))  

g5 <- ggplot(srilanka_weekly_seasonal, 
                     aes(x = week_of_year, y = cases, group = year, color = factor(year))) +
  geom_line(size=2) +
  facet_wrap(vars(district), scales = "free_y", ncol = 5) +
  scale_x_continuous(breaks = seq(1, 52, by = 4)) + 
  labs(
    x = "Week of Year",
    y = "Weekly Dengue Cases",
    color = "Year",
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 26, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),
    legend.title = element_text(size = 26, face = "bold"),
    legend.text  = element_text(size = 24),
    legend.key.size = unit(1.5, "cm")
  ) + scale_color_brewer(palette = "Paired")

g5
```


### Individual District Analysis

```{r}
seasonal <- function(dis_name) {

  dis <- srilanka_weekly_seasonal |>
    dplyr::filter(district == dis_name)

  ggplotly(
    ggplot(
      dis,
      aes(
        x = week_of_year,
        y = cases,
        group = year,
        color = factor(year)
      )
    ) +
      geom_line() +
      scale_x_continuous(breaks = seq(1, 52, by = 4)) +
      scale_color_viridis_d(option = "viridis") +
      labs(
        title = paste("Seasonal Plot -", dis_name),
        x = "Week of Year",
        y = "Weekly Dengue Cases",
        color = "Year"
      ) +
      theme_minimal()
  )
}
```

#### Column {.tabset}

##### Colombo

```{r}
seasonal("Colombo")
```

##### Gampaha

```{r}
seasonal("Gampaha")
```

##### Kalutara

```{r}
seasonal("Kalutara")
```

##### Kandy

```{r}
seasonal("Kandy")
```

##### Matale

```{r}
seasonal("Matale")
```

##### NuwaraEliya

```{r}
seasonal("NuwaraEliya")
```

##### Galle

```{r}
seasonal("Galle")
```

##### Hambanthota

```{r}
seasonal("Hambanthota")
```

##### Matara

```{r}
seasonal("Matara")
```

##### Jaffna

```{r}
seasonal("Jaffna")
```

##### Kilinochchi

```{r}
seasonal("Kilinochchi")
```

##### Mannar

```{r}
seasonal("Mannar")
```

##### Vavuniya

```{r}
seasonal("Vavuniya")
```

##### Mullaitivu

```{r}
seasonal("Mullaitivu")
```

##### Batticaloa

```{r}
seasonal("Batticaloa")
```

##### Ampara

```{r}
seasonal("Ampara")
```

##### Trincomalee

```{r}
seasonal("Trincomalee")
```

##### Kurunegala

```{r}
seasonal("Kurunegala")
```

##### Puttalam

```{r}
seasonal("Puttalam")
```

##### Anuradhapura

```{r}
seasonal("Anuradhapura")
```

##### Polonnaruwa

```{r}
seasonal("Polonnaruwa")
```

##### Badulla

```{r}
seasonal("Badulla")
```

##### Monaragala

```{r}
seasonal("Monaragala")
```

##### Ratnapura

```{r}
seasonal("Ratnapura")
```

##### Kegalle

```{r}
seasonal("Kegalle")
```

##### Kalmune

```{r}
seasonal("Kalmune")
```


# Dengue Patients Distribution

## Column {.tabset}

### Total Dengue Patients Distribution in 2025

```{r}

```

### Dengue Patients Distribution by Year and District

```{r}

```

### Distribution of Yearly Dengue Patients for last 20 Years

```{r}

```

