---
title: "Sri Lanka Dengue Dashboard"
format:
  dashboard:
    css: styles.css
editor: source
---

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(denguedatahub)
library(ceylon)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(here)
library(stringdist)
library(sf)
library(viridis)
library(cartogram)
library(ISOweek)
library(lubridate)
library(xts)
library(dygraphs)
library(crosstalk)
library(leaflet)
library(htmltools)
library(plotly)
library(shiny)
library(RColorBrewer)
library(readxl)
library(rnaturalearth)
library(rnaturalearthdata)
library(scales) # For percentage formatting
library(GGally)
```

```{r}
#| label: load-data
#| message: false
#| warning: false
data(srilanka_weekly_data)

weekly_2026 <- read_csv(here("data","sri_lanka_weekly_dengue_cases_2026.csv"))
weekly_2026$year        <- as.numeric(weekly_2026$year)
weekly_2026$week        <- as.numeric(weekly_2026$week)
weekly_2026$start.date  <- as.Date(weekly_2026$start.date)
weekly_2026$district    <- as.character(weekly_2026$district)
weekly_2026$cases       <- as.numeric(weekly_2026$cases)

Global <- read_excel("data/dengue-global-data.xlsx")
Global$date      <- as.Date(Global$date )
Global$cases     <- as.numeric(Global$cases)
Global$country   <- as.character(Global$country )
Global$deaths    <- as.numeric(Global$deaths)
```

```{r}
#| label: cleandata

weekly_cases_by_date <- weekly_2026 |>
  group_by(start.date) |>
  summarise(weekly_cases = sum(cases, na.rm = TRUE)) |>
  arrange(start.date)
```

# Overview

::: {.yellow-box}

**Latest Dengue Surveillance Summary**\
**Last Updated:** `r format(Sys.Date(), "%A, %B %d, %Y")` \| **Latest local data available for:** May 17, 2026

:::


## Row {height="20%"}

```{r}
#| content: valuebox
#| title: "Total Cases in 2026"
#| color: "#d95f02"
#| icon: "virus"

sum(weekly_2026$cases, na.rm = TRUE)
```

```{r}
#| content: valuebox
#| title: "Average Weekly Cases in 2026"
#| color: "#7570b3"
#| icon: "calendar-week"

round(sum(weekly_2026$cases, na.rm = TRUE)/20)
```

```{r}
#| content: valuebox
#| title: "Highest Risk District in 2026"
#| icon: "geo-alt"
#| color: "#1b9e77"

weekly_2026 |>
  group_by(district) |>
  summarise(total_cases = sum(cases, na.rm = TRUE)) |>
  arrange(desc(total_cases)) |>
  slice(1) |>
  pull(district)
```

## Row {height="65%"}

### Column {width="40%"}

```{r}
#| label: map1
#| echo: false
#| message: false
#| warning: false


district_df <- weekly_2026 |>
  group_by(district) |>       
  summarise(Count = sum(cases), .groups = "drop")  

data(district)
district_df <- district_df |>
  mutate(district_clean = toupper(trimws(district)))

district_sf <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

district_df <- district_df |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf$DISTRICT_clean[
      amatch(district_clean, district_sf$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()

district_df2 <- as.data.frame(district_df)

map_data <- left_join(
  district_sf,
  district_df2,
  by = c("DISTRICT_clean" = "DISTRICT_match")
)


g1 <- ggplotly(ggplot(map_data) +
  geom_sf(aes(fill = Count, text = paste('District:', district))) +
  labs(title = "Map for Total Dengue Cases in 2026") +  
  scale_fill_viridis() +
  theme_minimal()+
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(0, 0, 0, 0)
  ))

g1 <- g1 |>
  layout(
    title = list(
      text = "Map for Distribution of Total Dengue Cases in 2026",
      font = list(size = 16)   
    ))

g1

```

### Column {width="40%"}

```{r}
#| label: map2
#| echo: false
#| message: false
#| warning: false


map_data <- map_data |>
mutate(
risk = Count / population
)

map_data <- map_data |>
filter(!is.na(risk) & risk > 0)

dorl <- cartogram_dorling(map_data, weight = "population") |>
  st_as_sf()

g2 <- ggplotly(
  ggplot(dorl) +
    geom_sf(data = map_data) +
    geom_sf(aes(
      fill = Count,
      text = paste(
        "District:", DISTRICT_clean
      )
    )) +
    labs(title = "Map for Dengue Cases in 2026 Weighted by Population") +
    scale_fill_viridis()+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
)

g2 <- g2 |>
    layout(
    title = list(
       text = "Map for Dengue Cases in 2026 weighted by population",
       font = list(size = 15)   
     ))

g2
```

### Column {width="40%"}

```{r}
#| label: TS1
#| echo: false
#| message: false
#| warning: false

g3 <- ggplotly(ggplot(weekly_cases_by_date, aes(x = start.date, y = weekly_cases,)) +
      geom_line(size = 1, col = "#3b0f70") +           
      geom_point(size = 1, col = "#3b0f70") +
      scale_color_viridis_d() +
      scale_x_date(
    date_labels = "%b %d",   
    date_breaks = "2 weeks") +
      labs(
        title = "Weekly Dengue Cases in 2026",
        x = "Date (Month-Day)",
        y = "Cases"
      ) +
      theme_minimal(base_size = 14)
    )

g3 <- g3 |>
  layout(
    title = list(
      text = "Weekly Dengue Cases in 2026",
      font = list(size = 16)   
    ),

    xaxis = list(
      title = "Date (Month-Day)",
      titlefont = list(size = 14),   
      tickfont = list(size = 10),    
      tickangle = -60                  
    ),

    yaxis = list(
      title = "Cases",
      titlefont = list(size = 14),
      tickfont = list(size = 12),
      range = c(0, 2000)
    )
  )

g3
```

# District-wise Time Series Trends

## Column {.tabset}

### Overview of District Trends

```{r}
#| label: TS2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20

 
srilanka_weekly_df <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  )

g4 <- ggplot(srilanka_weekly_df, aes(x = week_start, y = cases)) +
  geom_line(color = "darkblue", linewidth = 0.8) +
  facet_wrap(vars(district), scales = "free_y", ncol = 4) + 
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    x = "Year",
    y = "Weekly dengue cases"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 26, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28)
  )

g4
```

### Explore Individual District Trends

```{r}
#| label: TS3
#| echo: false
#| warning: false
#| message: false


srilanka_weekly <- srilanka_weekly_data |>
  mutate(start.date = mdy(start.date)) |>
  arrange(start.date)

# --- Colombo ---
colombo_data <- srilanka_weekly |> filter(district == "Colombo")
colombo_xts  <- xts(colombo_data$cases, order.by = colombo_data$start.date)
colnames(colombo_xts) <- "Cases"

g_colombo <- dygraph(colombo_xts, main = "Colombo: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Gampaha ---
gampaha_data <- srilanka_weekly |> filter(district == "Gampaha")
gampaha_xts  <- xts(gampaha_data$cases, order.by = gampaha_data$start.date)
colnames(gampaha_xts) <- "Cases"

g_gampaha <- dygraph(gampaha_xts, main = "Gampaha: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalutara ---
kalutara_data <- srilanka_weekly |> filter(district == "Kalutara")
kalutara_xts  <- xts(kalutara_data$cases, order.by = kalutara_data$start.date)
colnames(kalutara_xts) <- "Cases"

g_kalutara <- dygraph(kalutara_xts, main = "Kalutara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kandy ---
kandy_data <- srilanka_weekly |> filter(district == "Kandy")
kandy_xts  <- xts(kandy_data$cases, order.by = kandy_data$start.date)
colnames(kandy_xts) <- "Cases"

g_kandy <- dygraph(kandy_xts, main = "Kandy: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |>
  dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matale ---
matale_data <- srilanka_weekly |> filter(district == "Matale")
matale_xts  <- xts(matale_data$cases, order.by = matale_data$start.date)
colnames(matale_xts) <- "Cases"

g_matale <- dygraph(matale_xts, main = "Matale: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Nuwara Eliya ---
nuwara_eliya_data <- srilanka_weekly |> filter(district == "NuwaraEliya")
nuwara_eliya_xts  <- xts(nuwara_eliya_data$cases, order.by = nuwara_eliya_data$start.date)
colnames(nuwara_eliya_xts) <- "Cases"

g_nuwara_eliya <- dygraph(nuwara_eliya_xts, main = "Nuwara Eliya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Galle ---
galle_data <- srilanka_weekly |> filter(district == "Galle")
galle_xts  <- xts(galle_data$cases, order.by = galle_data$start.date)
colnames(galle_xts) <- "Cases"

g_galle <- dygraph(galle_xts, main = "Galle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Matara ---
matara_data <- srilanka_weekly |> filter(district == "Matara")
matara_xts  <- xts(matara_data$cases, order.by = matara_data$start.date)
colnames(matara_xts) <- "Cases"

g_matara <- dygraph(matara_xts, main = "Matara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Hambantota ---
hambantota_data <- srilanka_weekly |> filter(district == "Hambanthota")
hambantota_xts  <- xts(hambantota_data$cases, order.by = hambantota_data$start.date)
colnames(hambantota_xts) <- "Cases"

g_hambantota <- dygraph(hambantota_xts, main = "Hambanthota: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Jaffna ---
jaffna_data <- srilanka_weekly |> filter(district == "Jaffna")
jaffna_xts  <- xts(jaffna_data$cases, order.by = jaffna_data$start.date)
colnames(jaffna_xts) <- "Cases"

g_jaffna <- dygraph(jaffna_xts, main = "Jaffna: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kilinochchi ---
kilinochchi_data <- srilanka_weekly |> filter(district == "Kilinochchi")
kilinochchi_xts  <- xts(kilinochchi_data$cases, order.by = kilinochchi_data$start.date)
colnames(kilinochchi_xts) <- "Cases"

g_kilinochchi <- dygraph(kilinochchi_xts, main = "Kilinochchi: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mannar ---
mannar_data <- srilanka_weekly |> filter(district == "Mannar")
mannar_xts  <- xts(mannar_data$cases, order.by = mannar_data$start.date)
colnames(mannar_xts) <- "Cases"

g_mannar <- dygraph(mannar_xts, main = "Mannar: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Vavuniya ---
vavuniya_data <- srilanka_weekly |> filter(district == "Vavuniya")
vavuniya_xts  <- xts(vavuniya_data$cases, order.by = vavuniya_data$start.date)
colnames(vavuniya_xts) <- "Cases"

g_vavuniya <- dygraph(vavuniya_xts, main = "Vavuniya: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Mullaitivu ---
mullaitivu_data <- srilanka_weekly |> filter(district == "Mullaitivu")
mullaitivu_xts  <- xts(mullaitivu_data$cases, order.by = mullaitivu_data$start.date)
colnames(mullaitivu_xts) <- "Cases"

g_mullaitivu <- dygraph(mullaitivu_xts, main = "Mullaitivu: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Batticaloa ---
batticaloa_data <- srilanka_weekly |> filter(district == "Batticaloa")
batticaloa_xts  <- xts(batticaloa_data$cases, order.by = batticaloa_data$start.date)
colnames(batticaloa_xts) <- "Cases"

g_batticaloa <- dygraph(batticaloa_xts, main = "Batticaloa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ampara ---
ampara_data <- srilanka_weekly |> filter(district == "Ampara")
ampara_xts  <- xts(ampara_data$cases, order.by = ampara_data$start.date)
colnames(ampara_xts) <- "Cases"

g_ampara <- dygraph(ampara_xts, main = "Ampara: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Trincomalee ---
trincomalee_data <- srilanka_weekly |> filter(district == "Trincomalee")
trincomalee_xts  <- xts(trincomalee_data$cases, order.by = trincomalee_data$start.date)
colnames(trincomalee_xts) <- "Cases"

g_trincomalee <- dygraph(trincomalee_xts, main = "Trincomalee: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kurunegala ---
kurunegala_data <- srilanka_weekly |> filter(district == "Kurunegala")
kurunegala_xts  <- xts(kurunegala_data$cases, order.by = kurunegala_data$start.date)
colnames(kurunegala_xts) <- "Cases"

g_kurunegala <- dygraph(kurunegala_xts, main = "Kurunegala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Puttalam ---
puttalam_data <- srilanka_weekly |> filter(district == "Puttalam")
puttalam_xts  <- xts(puttalam_data$cases, order.by = puttalam_data$start.date)
colnames(puttalam_xts) <- "Cases"

g_puttalam <- dygraph(puttalam_xts, main = "Puttalam: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Anuradhapura ---
anuradhapura_data <- srilanka_weekly |> filter(district == "Anuradhapura")
anuradhapura_xts  <- xts(anuradhapura_data$cases, order.by = anuradhapura_data$start.date)
colnames(anuradhapura_xts) <- "Cases"

g_anuradhapura <- dygraph(anuradhapura_xts, main = "Anuradhapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Polonnaruwa ---
polonnaruwa_data <- srilanka_weekly |> filter(district == "Polonnaruwa")
polonnaruwa_xts  <- xts(polonnaruwa_data$cases, order.by = polonnaruwa_data$start.date)
colnames(polonnaruwa_xts) <- "Cases"

g_polonnaruwa <- dygraph(polonnaruwa_xts, main = "Polonnaruwa: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Badulla ---
badulla_data <- srilanka_weekly |> filter(district == "Badulla")
badulla_xts  <- xts(badulla_data$cases, order.by = badulla_data$start.date)
colnames(badulla_xts) <- "Cases"

g_badulla <- dygraph(badulla_xts, main = "Badulla: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Moneragala ---
moneragala_data <- srilanka_weekly |> filter(district == "Monaragala")
moneragala_xts  <- xts(moneragala_data$cases, order.by = moneragala_data$start.date)
colnames(moneragala_xts) <- "Cases"

g_moneragala <- dygraph(moneragala_xts, main = "Monaragala: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Ratnapura ---
ratnapura_data <- srilanka_weekly |> filter(district == "Ratnapura")
ratnapura_xts  <- xts(ratnapura_data$cases, order.by = ratnapura_data$start.date)
colnames(ratnapura_xts) <- "Cases"

g_ratnapura <- dygraph(ratnapura_xts, main = "Ratnapura: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kegalle ---
kegalle_data <- srilanka_weekly |> filter(district == "Kegalle")
kegalle_xts  <- xts(kegalle_data$cases, order.by = kegalle_data$start.date)
colnames(kegalle_xts) <- "Cases"

g_kegalle <- dygraph(kegalle_xts, main = "Kegalle: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")

# --- Kalmunai ---
kalmunai_data <- srilanka_weekly |> filter(district == "Kalmune")
kalmunai_xts  <- xts(kalmunai_data$cases, order.by = kalmunai_data$start.date)
colnames(kalmunai_xts) <- "Cases"

g_kalmunai <- dygraph(kalmunai_xts, main = "Kalmune: Weekly Dengue Cases") |>
  dyRangeSelector(height = 40) |> dyOptions(strokeWidth = 2.5, drawGrid = TRUE, colors = "#1b9e77")


plot_containers <- tags$div(
  style = "width: 100%; margin-top: 15px;",
  
  tags$div(id = "plot-colombo", class = "district-plot", style = "display: block;", g_colombo),
  tags$div(id = "plot-gampaha", class = "district-plot", style = "display: none;", g_gampaha),
  tags$div(id = "plot-kalutara", class = "district-plot", style = "display: none;", g_kalutara),
  tags$div(id = "plot-kandy", class = "district-plot", style = "display: none;", g_kandy),
  tags$div(id = "plot-matale", class = "district-plot", style = "display: none;", g_matale),
  tags$div(id = "plot-nuwara-eliya", class = "district-plot", style = "display: none;", g_nuwara_eliya),
  tags$div(id = "plot-galle", class = "district-plot", style = "display: none;", g_galle),
  tags$div(id = "plot-matara", class = "district-plot", style = "display: none;", g_matara),
  tags$div(id = "plot-hambantota", class = "district-plot", style = "display: none;", g_hambantota),
  tags$div(id = "plot-jaffna", class = "district-plot", style = "display: none;", g_jaffna),
  tags$div(id = "plot-kilinochchi", class = "district-plot", style = "display: none;", g_kilinochchi),
  tags$div(id = "plot-mannar", class = "district-plot", style = "display: none;", g_mannar),
  tags$div(id = "plot-vavuniya", class = "district-plot", style = "display: none;", g_vavuniya),
  tags$div(id = "plot-mullaitivu", class = "district-plot", style = "display: none;", g_mullaitivu),
  tags$div(id = "plot-batticaloa", class = "district-plot", style = "display: none;", g_batticaloa),
  tags$div(id = "plot-ampara", class = "district-plot", style = "display: none;", g_ampara),
  tags$div(id = "plot-trincomalee", class = "district-plot", style = "display: none;", g_trincomalee),
  tags$div(id = "plot-kurunegala", class = "district-plot", style = "display: none;", g_kurunegala),
  tags$div(id = "plot-puttalam", class = "district-plot", style = "display: none;", g_puttalam),
  tags$div(id = "plot-anuradhapura", class = "district-plot", style = "display: none;", g_anuradhapura),
  tags$div(id = "plot-polonnaruwa", class = "district-plot", style = "display: none;", g_polonnaruwa),
  tags$div(id = "plot-badulla", class = "district-plot", style = "display: none;", g_badulla),
  tags$div(id = "plot-moneragala", class = "district-plot", style = "display: none;", g_moneragala),
  tags$div(id = "plot-ratnapura", class = "district-plot", style = "display: none;", g_ratnapura),
  tags$div(id = "plot-kegalle", class = "district-plot", style = "display: none;", g_kegalle),
  tags$div(id = "plot-kalmunai", class = "district-plot", style = "display: none;", g_kalmunai)
)


dropdown_selector <- tags$select(
  id = "custom-district-selector",
  class = "form-select",
  style = "width: 100%; max-width: 300px; padding: 8px; border-radius: 5px; border: 1px solid #ccc; font-size: 14px;",
  onchange = "
    // Hide all plot containers
    document.querySelectorAll('.district-plot').forEach(el => el.style.display = 'none');
    
    // Show only the selected container
    const targetedPlot = document.getElementById(this.value);
    if (targetedPlot) { targetedPlot.style.display = 'block'; }
    
    // Force dygraphs to redraw perfectly inside the unhidden space
    window.dispatchEvent(new Event('resize'));
  ",
  tags$option(value = "plot-colombo", "Colombo"),
  tags$option(value = "plot-gampaha", "Gampaha"),
  tags$option(value = "plot-kalutara", "Kalutara"),
  tags$option(value = "plot-kandy", "Kandy"),
  tags$option(value = "plot-matale", "Matale"),
  tags$option(value = "plot-nuwara-eliya", "Nuwara Eliya"),
  tags$option(value = "plot-galle", "Galle"),
  tags$option(value = "plot-matara", "Matara"),
  tags$option(value = "plot-hambantota", "Hambanthota"),
  tags$option(value = "plot-jaffna", "Jaffna"),
  tags$option(value = "plot-kilinochchi", "Kilinochchi"),
  tags$option(value = "plot-mannar", "Mannar"),
  tags$option(value = "plot-vavuniya", "Vavuniya"),
  tags$option(value = "plot-mullaitivu", "Mullaitivu"),
  tags$option(value = "plot-batticaloa", "Batticaloa"),
  tags$option(value = "plot-ampara", "Ampara"),
  tags$option(value = "plot-trincomalee", "Trincomalee"),
  tags$option(value = "plot-kurunegala", "Kurunegala"),
  tags$option(value = "plot-puttalam", "Puttalam"),
  tags$option(value = "plot-anuradhapura", "Anuradhapura"),
  tags$option(value = "plot-polonnaruwa", "Polonnaruwa"),
  tags$option(value = "plot-badulla", "Badulla"),
  tags$option(value = "plot-moneragala", "Monaragala"),
  tags$option(value = "plot-ratnapura", "Ratnapura"),
  tags$option(value = "plot-kegalle", "Kegalle"),
  tags$option(value = "plot-kalmunai", "Kalmune")
)


browsable(
  tags$div(
    style = "padding: 10px; background: #fff; border-radius: 5px;",
    tags$label(tags$strong("Select District:")),
    tags$div(style = "margin-top: 5px;", dropdown_selector),
    plot_containers
  )
)
```

# District-wise Seasonal Plots

## Column {.tabset}

### District Overview

```{r}
#| label: Seasonal1
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 20


srilanka_weekly_seasonal <- srilanka_weekly_data |>
  mutate(
    year_week = paste0(year, "-W", sprintf("%02d", week)),  
    week_start = ISOweek2date(paste0(year_week, "-1"))      
  ) |>
  filter(year_week >= 2020)


srilanka_weekly_seasonal <- srilanka_weekly_seasonal |>
  mutate(week_of_year = isoweek(week_start))  

g5 <- ggplot(srilanka_weekly_seasonal, 
                     aes(x = week_of_year, y = cases, group = year, color = factor(year))) +
  geom_line(size=2) +
  facet_wrap(vars(district), scales = "free_y", ncol = 5) +
  scale_x_continuous(breaks = seq(1, 52, by = 4)) + 
  labs(
    x = "Week of Year",
    y = "Weekly Dengue Cases",
    color = "Year",
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 24, face = "bold"),
    axis.text.y = element_text(size = 24, face = "bold"),
    strip.text = element_text(size = 24, face = "bold"),
    plot.title = element_text(size = 26, face = "bold", margin = margin(b = 10)),
    axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),
    legend.title = element_text(size = 26, face = "bold"),
    legend.text  = element_text(size = 24),
    legend.key.size = unit(1.5, "cm")
  ) + scale_color_brewer(palette = "Paired")

g5
```

### Individual District Analysis

```{r}
#| label: Seasonal2
#| echo: false
#| warning: false
#| message: false

seasonal <- function(dis_name) {

  dis <- srilanka_weekly_seasonal |>
    dplyr::filter(district == dis_name)

  ggplotly(
    ggplot(
      dis,
      aes(
        x = week_of_year,
        y = cases,
        group = year,
        color = factor(year)
      )
    ) +
      geom_line() +
      scale_x_continuous(breaks = seq(1, 52, by = 4)) +
      scale_color_viridis_d(option = "viridis") +
      labs(
        title = paste("Seasonal Plot -", dis_name),
        x = "Week of Year",
        y = "Weekly Dengue Cases",
        color = "Year"
      ) +
      theme_minimal()
  )
}
```

#### Column {.tabset}

##### Colombo

```{r}
seasonal("Colombo")
```

##### Gampaha

```{r}
seasonal("Gampaha")
```

##### Kalutara

```{r}
seasonal("Kalutara")
```

##### Kandy

```{r}
seasonal("Kandy")
```

##### Matale

```{r}
seasonal("Matale")
```

##### NuwaraEliya

```{r}
seasonal("NuwaraEliya")
```

##### Galle

```{r}
seasonal("Galle")
```

##### Hambanthota

```{r}
seasonal("Hambanthota")
```

##### Matara

```{r}
seasonal("Matara")
```

##### Jaffna

```{r}
seasonal("Jaffna")
```

##### Kilinochchi

```{r}
seasonal("Kilinochchi")
```

##### Mannar

```{r}
seasonal("Mannar")
```

##### Vavuniya

```{r}
seasonal("Vavuniya")
```

##### Mullaitivu

```{r}
seasonal("Mullaitivu")
```

##### Batticaloa

```{r}
seasonal("Batticaloa")
```

##### Ampara

```{r}
seasonal("Ampara")
```

##### Trincomalee

```{r}
seasonal("Trincomalee")
```

##### Kurunegala

```{r}
seasonal("Kurunegala")
```

##### Puttalam

```{r}
seasonal("Puttalam")
```

##### Anuradhapura

```{r}
seasonal("Anuradhapura")
```

##### Polonnaruwa

```{r}
seasonal("Polonnaruwa")
```

##### Badulla

```{r}
seasonal("Badulla")
```

##### Monaragala

```{r}
seasonal("Monaragala")
```

##### Ratnapura

```{r}
seasonal("Ratnapura")
```

##### Kegalle

```{r}
seasonal("Kegalle")
```

##### Kalmune

```{r}
seasonal("Kalmune")
```

# Dengue Patients Distribution

## Column {.tabset}

### Total Dengue Patients Distribution in 2025 (Last Year)

```{r}
#| label: Dis1
#| echo: false
#| warning: false
#| message: false

srilanka_2025 <- srilanka_weekly_data |>
  filter(year==2025) |> group_by(district) |>
  summarise(cases = sum(cases))

srilanka_2025 <- srilanka_2025 |>
  mutate(parents = "2025")

plot_ly(
    data = srilanka_2025,
     type="treemap",
     labels = ~ district,
     values = ~ cases,
    parents = ~ 2025,
    domain = list(column = 0),
    textinfo = "label+value+ percent parent"
  ) 

```

### Dengue Patients Distribution by Year and District

```{r}
#| label: Dis2
#| echo: false
#| warning: false
#| message: false
#| fig-width: 80
#| fig-height: 30

ann_df <- srilanka_weekly_data |>
  group_by(year, district) |>        
  summarise(annual_cases = sum(cases), .groups = "drop")  
data(district)
ann_df <- ann_df |>
  mutate(district_clean = toupper(trimws(district)))

dis_sf <- district |>
  mutate(DISTRICT_clean = toupper(trimws(DISTRICT)))

ann_df <- ann_df |>
  rowwise() |>
  mutate(
    DISTRICT_match = district_sf$DISTRICT_clean[
      amatch(district_clean, district_sf$DISTRICT_clean, maxDist = 3)
    ]
  ) |>
  ungroup()
ann_sf <- dis_sf |>
  left_join(ann_df, by = c("DISTRICT_clean" = "DISTRICT_match"))
ann_sf <- ann_sf |> filter(year > 2006)
ann_sf <- ann_sf |>
  mutate(
    cases_cat = cut(
      annual_cases,
      breaks = c(0, 30, 60, 100, 150, 250, 500, 1000, Inf),
      labels = c("0-30","31-60", "61-100", "101-150", "151-250",
                 "251-500", "501-1000", "Above 1000"),
      include.lowest = TRUE,
      right = TRUE
    )
  )


g6 <- ggplot(ann_sf) +
  geom_sf(aes(fill = cases_cat), color = "grey50", lwd = 0.2) +
  scale_fill_viridis_d(    
    option = "magma",
    direction = -1,
    na.value = "white",
    name = "Annual Dengue Cases"
  ) +
  facet_wrap(~year, ncol=10
             ) +
  theme_minimal() +
  coord_sf() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 26),
    strip.text = element_text(size = 28, face = "bold"),
    legend.title = element_text(size = 30, face = "bold"),
    legend.text  = element_text(size = 28),
    legend.key.size = unit(1.5, "cm"),
    plot.margin = margin(0, 0, 0, 0)
  )

g6

```

### Distribution of Yearly Dengue Patients for last 20 Years

```{r}
#| label: Dis3
#| echo: false
#| warning: false
#| message: false

ann_df <- ann_df %>%
  filter(year > 2006) %>%
  mutate(
    district = factor(district, levels = unique(district)),  # keep district order
    cases_cat = cut(
      annual_cases,
      breaks = c(0, 30, 60, 100, 150, 250, 500, 1000, Inf),
      labels = c("0-30","31-60", "61-100", "101-150", "151-250",
                 "251-500", "501-1000", "Above 1000"),
      include.lowest = TRUE
    )
  )
g7 <- ggplot(ann_df) +
  geom_tile(aes(
    x = year, 
    y = district, 
    fill = cases_cat,
    text = paste0("District: ", district,
                  "<br>Year: ", year,
                  "<br>Annual Cases: ", annual_cases)
  ), color = "white") +
  scale_fill_viridis_d(
    option = "magma",  
    direction = -1,
    na.value = "grey90",
    name = "Annual Dengue Cases"
  ) +
  theme_minimal() +
  scale_x_continuous(
    breaks = unique(ann_df$year)  
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 8)
  )
ggplotly(g7, tooltip = "text")

```

# Global Comparison

## Column {.tabset}

### Global Distribution of the Patient in 2025

```{r}
#| label: global1
#| echo: false
#| warning: false
#| message: false

Global <- Global |>
  mutate(year = year(date))

global_2025 <- Global |>
  filter(year == 2025) |>
  group_by(country) |>
  summarise(cases = sum(cases), deaths = sum(deaths)) |>
  arrange(desc(cases))

world <- ne_countries(scale = "medium", returnclass = "sf")

world2 <- world |>
  left_join(
    global_2025,
    by = c("name" = "country")
  )

world2 <- world2 |>
  mutate(
    cases_cat = cut(
      cases,
      breaks = c(-1, 0, 499, 4999, 49999, Inf),
      labels = c(
        "0",
        "1-499",
        "500-4,999",
        "5,000-49,999",
        "50,000+"
      ),
      include.lowest = TRUE
    )
  )

pal <- colorFactor(
  palette = c(
    "#F6EDB3",
    "#FDC777",
    "#F89C59",
    "#F15B4A",
    "#C91F5C"
  ),
  domain = world2$cases_cat,
  na.color = "#EEEEEE"
)

g8 <- leaflet(
  world2,
  options = leafletOptions(
    worldCopyJump = FALSE,
    minZoom = 2
  )) |>
  
  addProviderTiles(
    "CartoDB.Positron",
    options = providerTileOptions(noWrap = TRUE)
  )  |>

     addPolygons(
       fillColor = ~pal(cases_cat),
       color = "#cccccc",
       weight = 0.5,
       fillOpacity = 0.9,

    popup = ~paste0(
      "<b>", name, "</b><br>",
      "Cases: ", cases
    )
  ) |>

  addLegend(
    pal = pal,
    values = ~cases_cat,
    title = "Total Cases",
    position = "bottomright"
  )

g8 <- g8 |>
  htmlwidgets::onRender("
    function(el,x){
      this.invalidateSize();
    }
  ")

g8
```

### Sri Lanka vs. Top 10 Countries Worldwide (2025)

#### Column {width="50%"}

```{r}
#| label: global2
#| echo: false
#| warning: false
#| message: false

global_top10 <- global_2025 |>
  arrange(desc(cases)) |>
  slice_head(n = 10)

sri_lanka <- global_2025 |>
  filter(country == "Sri Lanka")

global_top10 <- bind_rows(global_top10, sri_lanka) |>
  distinct(country, .keep_all = TRUE)

global_top10$country <- factor(
  global_top10$country,
  levels = rev(global_top10$country)
)

global_top10 <- global_top10 |> 
  pivot_longer( cols = c(cases, deaths), names_to = "type", values_to = "count" ) |>
  mutate(
    count = replace_na(count, 0)
  )


g9 <- plot_ly(
  global_top10,
  x = ~count,
  y = ~country,
  type = "bar",
  color = ~type,
  colors = c(
    "cases" = "steelblue",
    "deaths" = "firebrick"
  )
) |>
  layout(
    title = list(
      text = "Total Dengue Cases and Deaths by Country (2025)",
      font = list(size = 16)
    ),
    xaxis = list(title = "Count"),
    yaxis = list(title = list(text = "Country", standoff = 10)),
    barmode = "stack",
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5
    )
  )

g9
```

#### Column {width="50%"}

```{r}
#| label: global3
#| echo: false
#| warning: false
#| message: false


global_top10_rate <- global_2025 |>
  arrange(desc(cases)) |>
  slice_head(n = 10)

sri_lanka <- global_2025 |>
  filter(country == "Sri Lanka")

global_top10_rate <- bind_rows(global_top10_rate, sri_lanka) |>
  distinct(country, .keep_all = TRUE)

global_top10_rate <- global_top10_rate |>
  mutate(
    deaths = replace_na(deaths, 0)
  ) |>
  mutate(
    rate = (deaths/cases) * 100
  )

global_top10_rate$country <- factor(
  global_top10_rate$country,
  levels = rev(global_top10_rate$country)
)

g10 <- plot_ly(
  global_top10_rate,
  x = ~rate,
  y = ~reorder(country, rate),
  type = "bar",
  marker = list(color = "#8B0000"),
  name = "Death Rate",
  showlegend = FALSE
) |>
  layout(
    title = list(text = "Dengue Death Rate by Country (2025)", size = 16),
    xaxis = list(title = 'Death Rate (%)'),
    yaxis = list(
     title = list(text = "Country", standoff = 10)
    )
  )

g10
```

### Sri Lanka vs. South Eastern Asian Countries-WHO Region (2025)

#### Column {width="50%"}

```{r}
global_2025_asia <- Global |>
  filter(year == 2025 & who_region == "SEAR") |>
  group_by(country) |>
  summarise(cases = sum(cases), deaths = sum(deaths)) |>
  arrange(desc(cases))

global_2025_asia$country <- factor(
  global_2025_asia$country,
  levels = rev(global_2025_asia$country)
)

global_2025_asia_new <- global_2025_asia |> 
  pivot_longer( cols = c(cases, deaths), names_to = "type", values_to = "count" ) |>
  mutate(
    count = replace_na(count, 0)
  )

g11 <- plot_ly(
  global_2025_asia_new,
  x = ~count,
  y = ~country,
  type = "bar",
  color = ~type,
  colors = c(
    "cases" = "steelblue",
    "deaths" = "firebrick"
  )
) |>
  layout(
    title = list(
      text = "Total Dengue Cases and Deaths by Country (2025)",
      font = list(size = 16)
    ),
    xaxis = list(title = "Count"),
    yaxis = list(title = list(text = "Country", standoff = 10)),
    barmode = "stack",
    legend = list(
      orientation = "h",
      xanchor = "center",
      x = 0.5
    )
  )

g11

```

#### Column {width="50%"}

```{r}
global_2025_asia_rate <- global_2025_asia |>
  mutate(
    deaths = replace_na(deaths, 0)
  ) |>
  mutate(
    rate = (deaths/cases) * 100
  )

global_2025_asia_rate$country <- factor(
  global_2025_asia_rate$country,
  levels = rev(global_2025_asia_rate$country)
)

g12 <- plot_ly(
  global_2025_asia_rate,
  x = ~rate,
  y = ~reorder(country, rate),
  type = "bar",
  marker = list(color = "#8B0000"),
  name = "Death Rate",
  showlegend = FALSE
) |>
  layout(
    title = list(text = "Dengue Death Rate by Country (2025)", size = 16),
    xaxis = list(title = 'Death Rate (%)'),
    yaxis = list(
     title = list(text = "Country", standoff = 10)
    )
  )

g12
```

# About
